---
title: "00_differentialMethylationSequencing"
author: "Caitlin Page"
date: "2024-10-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Data
The data used in this analysis is B cell types and NK cell types from WGBS, 
under the hg19 genome annotation. [GSE186458](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE186458)

* B cell type: 3 samples
* NK cell type: 3 samples

## Packages used

```{r}
library(DMRcate)
library(bsseq)
library(edgeR)
library(limma)
library(BiocParallel)
library(BSgenome.Hsapiens.UCSC.hg19)
library(Biostrings)

library(plyranges)
library(dplyr)
library(ggplot2)
library(ggVennDiagram)
library(ComplexUpset)
```

## Obtain counts
```{r, eval=FALSE}
# get the positions of cg sites
seq_names <- seqnames(BSgenome.Hsapiens.UCSC.hg19)[1:25]
anno_seq <- lapply(seq_names, function(x) {
  cbind(data.frame(matchPattern("CG", BSgenome.Hsapiens.UCSC.hg19[[x]])),
        seqnames = x)}) %>%
  dplyr::bind_rows()
anno_seq <- anno_seq %>% mutate(pos = paste0(seqnames, "-", start)) %>%
  relocate(pos, seqnames)

# NK counts
fname <- "../data/wgbs/GSM5652299_Blood-NK-Z000000TM.beta"
N <- file.info(fname)$size
beta_nk <- data.frame(matrix(readBin(fname, "integer", N, size = 1, signed = FALSE), N / 2, 2, byrow=TRUE))
fname <- "../data/wgbs/GSM5652300_Blood-NK-Z000000U1.beta"
N <- file.info(fname)$size
beta_nk <- cbind(beta_nk, data.frame(matrix(readBin(fname, "integer", N, size = 1, signed = FALSE), N / 2, 2, byrow=TRUE)))
fname <- "../data/wgbs/GSM5652301_Blood-NK-Z000000UF.beta"
N <- file.info(fname)$size
beta_nk <- cbind(beta_nk, data.frame(matrix(readBin(fname, "integer", N, size = 1, signed = FALSE), N / 2, 2, byrow=TRUE)))
colnames(beta_nk) <- c("M_TM", "C_TM", "M_U1", "C_U1", "M_UF", "C_UF")
beta_nk <- cbind(anno_seq[,1:4], beta_nk)

# B counts
fname <- "../data/wgbs/GSM5652316_Blood-B-Z000000TX.beta"
N <- file.info(fname)$size
beta_b <- data.frame(matrix(readBin(fname, "integer", N, size = 1, signed = FALSE), N / 2, 2, byrow=TRUE))
fname <- "../data/wgbs/GSM5652317_Blood-B-Z000000UB.beta"
N <- file.info(fname)$size
beta_b <- cbind(beta_b, data.frame(matrix(readBin(fname, "integer", N, size = 1, signed = FALSE), N / 2, 2, byrow=TRUE)))
fname <- "../data/wgbs/GSM5652318_Blood-B-Z000000UR.beta"
N <- file.info(fname)$size
beta_b <- cbind(beta_b, data.frame(matrix(readBin(fname, "integer", N, size = 1, signed = FALSE), N / 2, 2, byrow=TRUE)))
colnames(beta_b) <- c("M_TX", "C_TX", "M_UB", "C_UB", "M_UR", "C_UR")
beta_b <- cbind(anno_seq[,1:4], beta_b)

# add unmeth counts
beta_nk <- beta_nk %>% mutate(Un_TM = C_TM - M_TM, Un_U1 = C_U1 - M_U1, Un_UF = C_UF - M_UF)
beta_b <- beta_b %>% mutate(Un_TX = C_TX - M_TX, Un_UB = C_UB - M_UB, Un_UR = C_UR - M_UR)

# combine
colnames(beta_b)[5:13] <- paste0("b_", colnames(beta_b)[5:13])
colnames(beta_nk)[5:13] <- paste0("nk_", colnames(beta_nk)[5:13])

wgbs_counts <- cbind(beta_nk, beta_b[,5:13])

saveRDS(wgbs_counts, "../output/wgbs_counts.rds")
```


## Differential Methylation Analysis

# bsseq pre-processing: bsseq and bsmooth objects
```{r eval=FALSE}
## set up
wgbs_counts <- wgbs_counts %>% filter(!seqnames %in% c("chrX", "chrY", "chrM"))
m_df <- wgbs_counts[,c(5,7,9,14,16,18)]
colnames(m_df) <- colnames(m_df) %>% gsub("_M", "", .)

coverage_df <- wgbs_counts[,c(6,8,10,15,17,19)]
colnames(coverage_df) <- colnames(coverage_df) %>% gsub("_C", "", .)

sample_stuff <- cbind(Type = c("nk", "nk", "nk", "b", "b", "b")) %>% data.frame()
rownames(sample_stuff) <- colnames(m_df)

## make BSseq object
bseq_nk_vs_b_all <- BSseq(M = as.matrix(m_df),
                          Cov = as.matrix(coverage_df), pData = sample_stuff,
                          chr = beta_all$seqnames, pos = as.numeric(beta_all$start),
                          sampleNames = colnames(m_df))

## BSmooth by sample - so it doesn't crash
bsmooth1 <- BSmooth(BSseq = bseq_nk_vs_b_all[,1], BPPARAM = MulticoreParam(workers = 6), verbose = TRUE)

bsmooth2 <- BSmooth(BSseq = bseq_nk_vs_b_all[,2], BPPARAM = MulticoreParam(workers = 6), verbose = TRUE)
bsmooth3 <- BSmooth(BSseq = bseq_nk_vs_b_all[,3], BPPARAM = MulticoreParam(workers = 6), verbose = TRUE)

bsmooth4 <- BSmooth(BSseq = bseq_nk_vs_b_all[,4], BPPARAM = MulticoreParam(workers = 6), verbose = TRUE)
bsmooth5 <- BSmooth(BSseq = bseq_nk_vs_b_all[,5], BPPARAM = MulticoreParam(workers = 6), verbose = TRUE)
bsmooth6 <- BSmooth(BSseq = bseq_nk_vs_b_all[,6], BPPARAM = MulticoreParam(workers = 6), verbose = TRUE)

bsmooth_obj <- combine(bsmooth1, bsmooth2, bsmooth3, bsmooth4, bsmooth5, bsmooth6)
## filter
bseq.cov <- getCoverage(bsmooth_obj)
keep <- which(rowSums(bseq.cov[, bsmooth_obj$Type == "nk"] >= 2) >= 2 &
                rowSums(bseq.cov[, bsmooth_obj$Type == "b"] >= 2) >= 2)
length(keep)

bsmooth_obj <- bsmooth_obj[keep,]
saveRDS(bsmooth_obj, "../output/bsmooth_obj.rds")
```


# bsseq
```{r eval=FALSE}
## testing
bsseq.tstat <- BSmooth.tstat(bsmooth_obj, group1 = c("nk_TM", "nk_U1", "nk_UF"),
                            group2 = c("b_TX", "b_UB", "b_UR"),
                            estimate.var = "group2",
                            local.correct = TRUE,
                            verbose = TRUE, mc.cores = 6)

bsseq_res <- bsseq.tstat@stats %>% data.frame() %>%
  mutate(seqnames = data.frame(bsseq.tstat@gr)$seqnames,
         start = data.frame(bsseq.tstat@gr)$start,
         end = start, position = paste0(seqnames, "-", start)) %>%
  relocate(position)
saveRDS(bsseq_res, "../output/bsseq_res.rds")

## dmrs
bsseq_dmrs <- dmrFinder(bseq.tstat)
# filter the dmrs
bsseq_dmrs <- subset(bsseq_dmrs, n >= 3 & abs(meanDiff) >= 0.1)
colnames(bsseq_dmrs)[1] <- "seqnames"
bsseq_dmrs <- bsseq_dmrs %>%
  mutate(position = paste0(seqnames, "-", start)) %>% relocate(position)
bsseq_dmrs$rank <- 1:nrow(bsseq_dmrs)
saveRDS(bsseq_dmrs, "../output/bsseq_dmrs.rds")
```

```{r include=FALSE}
bsseq_res <- readRDS("../output/bsseq_res.rds")
bsseq_dmrs <- readRDS("../output/bsseq_dmrs.rds")
```

```{r}
nrow(bsseq_res)
head(bsseq_res)
```

```{r}
nrow(bsseq_dmrs)
head(bsseq_dmrs)
```

# DMRcate

* takes in smoothed samples from bsseq
* also has the same filtering recommendations as bsseq
```{r eval=FALSE}
bsmooth_obj

tissue <- factor(c("nk", "nk", "nk", "b", "b", "b"), levels = c("nk", "b"))
tissue
#Regular matrix design
design <- model.matrix(~tissue)
design

methdesign <- modelMatrixMeth(design)
methdesign

seq_annot <- sequencing.annotate(bsmooth_obj, methdesign, all.cov = TRUE, 
                                 coef = 8, fdr=0.05)
#saveRDS(seq_annot, "../output/dmrcate_seq_anno.rds")

seq_dmrcate.res <- dmrcate(seq_annot, C=2, min.cpgs = 5)

dmrcate_seq_dmr <- extractRanges(seq_dmrcate.res, genome = "hg19")
#saveRDS(dmrcate_seq_dmr, "../output/dmrcate_seq_dmr.rds")
```

```{r include=FALSE}
dmrcate_seq_anno <- readRDS("../output/dmrcate_seq_anno.rds")
dmrcate_seq_dmr <- readRDS("../output/dmrcate_seq_dmr.rds") %>% data.frame()
```

```{r}
nrow(dmrcate_seq_anno)
head(dmrcate_seq_anno)
```

```{r}
nrow(dmrcate_seq_dmr)
head(dmrcate_seq_dmr)
```


# dmrseq
* Uses bsseq object as input
* dmrseq failed.
```{r eval=FALSE}
bseq_nk_vs_b_all

loci.idx <- which(DelayedMatrixStats::rowSums2(getCoverage(bseq_nk_vs_b_all, type="Cov")==0) == 0)
sample.idx <- which(pData(bseq_nk_vs_b_all)$Type %in% c("nk", "b"))

bs.filtered <- bseq_nk_vs_b_all[loci.idx, sample.idx]

testCovariate <- "Type"
regions <- dmrseq(bs=bs.filtered,
                  cutoff = 0.05,
                  testCovariate=testCovariate) # failed with error
```

# DSS
* Uses bsseq object as input
* DMLtest failed to finish. Unclear as to why, 
  expected to be too computationally intensive
```{r eval=FALSE}
bseq_nk_vs_b_all
dss_dmlTest_all <- DMLtest(bseq_nk_vs_b_all,
                                  group1=c("nk_TM", "nk_U1", "nk_UF"),
                                  group2 = c("b_TX", "b_UB", "b_UR"),
                               smoothing = TRUE, BPPARAM = BatchtoolsParam()) #failed
dss_dmrs_pos_filt <- callDMR(dss_dmlTest_nk_vs_b_pos_filt, p.threshold=0.05)
dss_dmrs <- callDMR(dss_dmlTest_nk_vs_b, p.threshold=0.05)
```


# edgeR

```{r eval=FALSE}
# pre-processing
edge_matrix <- wgbs_counts[,c(5,11,7,12,9,13,14,20,16,21,18,22)]

colnames(edge_matrix) <- c(
  "NK1.Me", "NK1.Un", "NK2.Me", "NK2.Un", "NK3.Me", "NK3.Un",
  "B1.Me", "B1.Un", "B2.Me", "B2.Un", "B3.Me", "B3.Un")
rownames(edge_matrix) <- wgbs_counts$pos

edge_matrix <- edge_matrix[!grepl("chrX", rownames(edge_matrix)),]
edge_matrix <- edge_matrix[!grepl("chrY", rownames(edge_matrix)),]

sample <- gl(6,2,12)
methylation <- gl(2,1,12, labels=c("Me","Un"))
condition <- gl(2,6,12, labels=c("NK", "B"))
edge_sample_info <- cbind(sample, methylation, condition)
rownames(edge_sample_info) <- colnames(edge_matrix)
data.frame(edge_sample_info)


edge_dge <- DGEList(as.matrix(edge_matrix), samples = data.frame(edge_sample_info),
                    group = c(
                      "NK", "NK", "NK", "NK", "NK", "NK",
                      "B", "B", "B", "B", "B", "B"))

edge_coverage <- edge_dge$counts[, c(1,3,5,7,9,11)] + edge_dge$counts[, c(2,4,6,8,10,12)]
head(edge_coverage)
keep <- rowSums(edge_coverage >= 8) == 6

edge_dge <- edge_dge[keep,,keep.lib.sizes=FALSE]

edge_dge$samples

edge_total_lib_size <- filter(edge_dge$samples, methylation==1)$lib.size +
  filter(edge_dge$samples, methylation==2)$lib.size

edge_dge$samples$lib.size <- rep(edge_total_lib_size, each=2)
edge_dge$samples

# testing
edge_samples <- cbind(Group = c("NK", "NK", "NK", "B", "B", "B")) %>% data.frame()
rownames(edge_samples) <- c("NK1", "NK2", "NK3", "B1", "B2", "B3")
edge_samples <- edge_samples %>% mutate(Group = as.factor(Group))

edge_design <- model.matrix(~0 + Group, data = edge_samples)
colnames(edge_design) <- c("B", "NK")

edge_design <- modelMatrixMeth(edge_design)

edge_dge <- estimateDisp(edge_dge, edge_design, trend="none")

edge_fit <- glmFit(edge_dge, edge_design)
edge_contrasts <- makeContrasts(NK - B,
                                levels=edge_design)
edge_lrt <- glmLRT(edge_fit, contrast=edge_contrasts)

edge_res <- topTags(edge_lrt, n = Inf)

edge_res <- edge_res$table %>% mutate(signif.05 = ifelse(FDR <= 0.05, TRUE, FALSE))

#saveRDS(edge_res, "../output/edge_res.rds")
```
```{r include=FALSE}
edge_res <- readRDS("../output/seq_edge_res.rds")
```
```{r}
nrow(edge_res)
head(edge_res)
```




