---
title: "01_exploreBiasByBins"
author: "Caitlin Page"
date: "2024-09-05"
output: workflowr::wflow_html
editor_options: 
  chunk_output_type: inline
---
chunk_output_type: console
## Introduction

```{r}
library(plyranges)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ChIPseqSpikeInFree)
library(biovizBase)
library(BSgenome.Hsapiens.UCSC.hg19)
```

```{r}
wgbs_counts <- readRDS("output/wgbs_counts.rds")
bsseq_res <- readRDS("output/bsseq_res.rds")
bsseq_dmrs <- readRDS("output/bsseq_dmrs.rds")
```

# What bin size to choose?

```{r}
bsseq_dmrs %>%
  ggplot(aes(x = width, colour = seqnames)) +
  geom_density()
```

* dmrs are mostly < 2000 bp

```{r shared-code}
source("code/function_binBiasPlot.R")
plotBiasGenomeBins(wgbs_counts, bsseq_dmrs, bin_size = 2000)
```

* explore other bin sizes

```{r}
plotBiasGenomeBins(wgbs_counts, bsseq_dmrs, bin_size = 500)
```

```{r}
plotBiasGenomeBins(wgbs_counts, bsseq_dmrs, bin_size = 1000)
```

```{r}
plotBiasGenomeBins(wgbs_counts, bsseq_dmrs, bin_size = 5000)
```

* prefer 2000 bp bin length

```{r}
genome_bins <- plotBiasGenomeBins(wgbs_counts, bsseq_dmrs, bin_size = 2000)
```

Questions:
* why are some bins with high cgs have no dmrs?

# Strange proportions in big bins
Possible explanations
* telomeres - strange things at the starts and the ends of chromosomes
Test: plot the first and last 20 bins per chromosome - are these the bins with high cg numbers?

* tail
```{r}
genome_bins[[1]] %>% group_by(seqnames) %>% do(tail(., n = 20)) %>% mutate(n = 1:n()) %>%
  ggplot(aes(x = n, y = num_cg, colour = seqnames)) +
  geom_point()
```
* head
```{r}
genome_bins[[1]] %>% group_by(seqnames) %>% do(head(., n = 20)) %>% mutate(n = 1:n()) %>%
  ggplot(aes(x = n, y = num_cg, colour = seqnames)) +
  geom_point()
```

Result: ends of chromosomes are not the source of high cg bins

## GC content?

```{r}
genome_bins <- cbind(genome_bins[[1]], biovizBase::GCcontent(BSgenome.Hsapiens.UCSC.hg19, as_granges(genome_bins[[1]]))) %>% data.frame()
```

```{r}
genome_bins %>%
  ggplot(aes(x = num_cg, y = C.G, colour = has_dmr)) +
  geom_point(alpha = 0.3)
```

```{r}
genome_bins %>%
  ggplot(aes(x = C.G)) +
  geom_density()

genome_bins %>%
  ggplot(aes(x = C.G, colour = has_dmr)) +
  geom_density()
```

```{r}
cor.test(genome_bins$num_cg, genome_bins$C.G)
```

GC content is related to coverage

## Coverage in bins
```{r}
counts_per_bin <- find_overlaps(as_granges(genome_bins), as_granges(wgbs_counts)) %>% data.frame()
```

```{r}
counts_per_bin_sum <- counts_per_bin %>% group_by(num) %>% mutate(sum_nk_TM = sum(nk_C_TM), sum_nk_U1 = sum(nk_C_U1), sum_nk_UF = sum(nk_C_UF),
                                            sum_b_TX = sum(b_C_TX), sum_b_UB = sum(b_C_UB), sum_b_UR = sum(b_C_UR)) %>% distinct(bin_pos, num, sum_nk_TM, sum_nk_U1, sum_nk_UF, sum_b_TX, sum_b_UB, sum_b_UR)
counts_per_bin_sum$av_nk <- rowMeans(counts_per_bin_sum[,3:5])
counts_per_bin_sum$av_b <- rowMeans(counts_per_bin_sum[,6:8])
counts_per_bin_sum <- counts_per_bin_sum %>% ungroup() %>% data.frame()
```

```{r}
genome_bins <- genome_bins %>% mutate(av_nk = counts_per_bin_sum[match(.$bin_pos, counts_per_bin_sum$bin_pos), "av_nk"],
                                      av_b = counts_per_bin_sum[match(.$bin_pos, counts_per_bin_sum$bin_pos), "av_b"])
genome_bins[is.na(genome_bins)] <- 0
genome_bins
```

```{r}
genome_bins %>%
  ggplot(aes(y = av_nk, x = C.G)) +
  geom_point(colour = "red", alpha = 0.3) +
  geom_point(aes(y = av_b, x = C.G), colour = "black", alpha = 0.3)
```
```{r}
genome_bins %>%
  ggplot(aes(y = log2(av_nk), x = C.G)) +
  geom_point(colour = "red", alpha = 0.3) +
  geom_point(aes(y = log2(av_b), x = C.G), colour = "black", alpha = 0.3)
```

```{r}
genome_bins %>% mutate(num_cg_gr_200 = ifelse(num_cg > 200, TRUE, FALSE)) %>%
  ggplot(aes(y = av_nk, colour = num_cg_gr_200)) +
  geom_boxplot()

genome_bins %>% mutate(num_cg_gr_200 = ifelse(num_cg > 200, TRUE, FALSE)) %>%
  ggplot(aes(x = av_nk, colour = num_cg_gr_200)) +
  geom_density()
```

```{r}
genome_bins[1:5000,] %>%
  ggplot(aes(x = av_nk, y = av_b)) +
  geom_point()
```
```{r}
cor.test(genome_bins$av_nk, genome_bins$av_b)
```

## how common are these bins anyway?
```{r}
genome_bins %>%
  ggplot(aes(x = num_cg)) +
  geom_density()
genome_bins %>%
  ggplot(aes(x = num_cg, colour = has_dmr)) +
  geom_density()
```

