---
title: "02_exploreBiasGenes"
author: "Caitlin Page"
date: "2024-09-05"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

```{r}
library(plyranges)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ExperimentHub)
library(BSgenome.Hsapiens.UCSC.hg19)
```
```{r}
library(AnnotationHub)
ah <- AnnotationHub()
```
```{r}
ah$genome
query(query(query(ah, "hg19"), "Homo sapiens"), "Ensembl genes")
```
```{r}
ah_a <- ah[["AH5046"]]
ah_a[1:5,] %>% data.frame()
```
```{r}
query(query(query(ah, "hg19"), "Homo sapiens"), "UCSC genes")
ah[["AH5036"]][1:5,] %>% data.frame()
```
- where is the TSS hiding??
- dont want to have to need biomart

```{r}
#query(query(query(query(ah, "hg19"), "Homo sapiens"), "GRanges"), "GENCODE")
```
- ensembl, ucsc, and gencode don't give tss coords
```{r}
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
TxDb.Hsapiens.UCSC.hg19.knownGene
```
```{r}
genes(TxDb.Hsapiens.UCSC.hg19.knownGene)
```
- maybe like a combination of trancripts and genes?
- but that seems terrible


```{r}
eh <- ExperimentHub()
eh_genes <- eh[["EH3132"]]
eh_genes
```
```{r}
eh["EH3132"]
```

```{r}
eh$
head(unique(eh$species))
```
```{r}
query(eh, "Homo sapiens")
```

```{r}
wgbs_counts <- readRDS("output/wgbs_counts.rds")
bsseq_res <- readRDS("output/bsseq_res.rds")
bsseq_dmrs <- readRDS("output/bsseq_dmrs.rds")
```


# overlap genes and dmrs

* Distribution of gene widths
```{r}
eh_genes %>% data.frame() %>%
  ggplot(aes(x = width)) +
  geom_density()
eh_genes %>% data.frame() %>%
  ggplot(aes(x = log2(width))) +
  geom_density()
```


```{r}
find_overlaps(as_granges(bsseq_dmrs), eh_genes) %>% data.frame()
```
```{r}
bsseq_dmrs <- bsseq_dmrs %>% mutate(rank = 1:n())
find_overlaps(as_granges(bsseq_dmrs), eh_genes) %>% data.frame() %>% group_by(rank) %>% summarise(n_genes =n())
find_overlaps(as_granges(bsseq_dmrs), eh_genes) %>% data.frame() %>% group_by(rank) %>% summarise(n_genes =n()) %>%
  ungroup() %>% group_by(n_genes) %>% summarise(n())

find_overlaps(as_granges(bsseq_dmrs), eh_genes) %>% data.frame() %>% group_by(symbol) %>% summarise(n_dmr =n())
find_overlaps(as_granges(bsseq_dmrs), eh_genes) %>% data.frame() %>% group_by(symbol) %>% summarise(n_dmr =n()) %>%
  ungroup() %>% group_by(n_dmr) %>% summarise(n())
```
- ok yeah this is why i wanted to know
- because we have 1 dmr that overlaps with 21 genes
- and we have 1 gene that overlaps with 95 dmr

```{r}
overlap_gene_dmr <- find_overlaps(as_granges(bsseq_dmrs), eh_genes) %>% data.frame()
```

```{r}
eh_genes %>% data.frame() %>% mutate(dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     dmr = ifelse(!is.na(dmr), TRUE, FALSE)) %>%
  group_by(width, dmr) %>% summarise(n= n()) %>% group_by(width) %>% mutate(sum = sum(n), prop = n/sum, prop = ifelse(dmr == FALSE, 0, prop), num = n()) %>%
  mutate(bseq_dmr = ifelse(num == 1, TRUE, dmr)) %>% filter(dmr == TRUE) %>%
  ggplot(aes(x = width, y = prop)) +
  geom_point(alpha = 0.3)
```

```{r}
eh_genes %>% data.frame() %>% mutate(dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     dmr = ifelse(!is.na(dmr), TRUE, FALSE)) %>%
  group_by(width, dmr) %>% summarise(n= n()) %>% group_by(width) %>% mutate(sum = sum(n), prop = n/sum, prop = ifelse(dmr == FALSE, 0, prop), num = n()) %>%
  mutate(bseq_dmr = ifelse(num == 1, TRUE, dmr)) %>% filter(dmr == TRUE) %>%
  ggplot(aes(x = log2(width), y = prop)) +
  geom_point(alpha = 0.3)
```

```{r}
eh_genes <- eh_genes %>% data.frame() %>% mutate(num = 1:n())
eh_genes <- eh_genes %>% mutate(bin_pos = paste0(seqnames, "-", start))
counts_per_gene <- find_overlaps(as_granges(eh_genes), as_granges(wgbs_counts)) %>% data.frame()
#eh_genes <- eh_genes %>% mutate(bin_pos = paste0(seqnames, "-", start))
cg_per_bin <- counts_per_gene %>% group_by(num, bin_pos) %>% summarise(n_region = n())
cg_per_bin <- cg_per_bin %>% ungroup() %>% data.frame()

eh_genes <- eh_genes %>% mutate(num_cg = cg_per_bin[match(.$num, cg_per_bin$num), "n_region"])
eh_genes[is.na(eh_genes)] <- 0

```

```{r}
eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>% group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>%
    group_by(num_cg, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_cg) %>%
    distinct(num_cg, num_bins_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg) %>%
    .[order(.$num_cg),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = num_cg, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```
```{r}
#library(goseq)
```

- not sure if this looks right
```{r}
eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>% group_by(num_cg) %>% 
  group_by(num_cg, has_dmr) %>% summarise(num_per_dmr = n()) %>% group_by(num_cg) %>% mutate(prop_dmr = num_per_dmr/sum(num_per_dmr)) %>% filter(has_dmr == TRUE)
eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>% group_by(num_cg) %>% 
  group_by(num_cg, has_dmr) %>% summarise(num_per_dmr = n()) %>% group_by(num_cg) %>% mutate(prop_dmr = num_per_dmr/sum(num_per_dmr)) %>% filter(has_dmr == TRUE) %>%
  ggplot(aes(x = num_cg, y = prop_dmr)) +
  geom_point(alpha = 0.4)
```

```{r}
eh_genes %>%
  ggplot(aes(x = num_cg, y = width)) +
  geom_point(alpha = 0.4)
```

# how many genes actually have a dmr?
```{r}
eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>%
  group_by(has_dmr) %>% summarise(n=n()) %>% ungroup() %>% mutate(prop = n/sum(n))
```

```{r}
eh_genes <- eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE))
```

# what prop of dmrs overlap to gene?

```{r}
bsseq_dmrs %>% mutate(overlap_gene = overlap_gene_dmr[match(.$rank, overlap_gene_dmr$rank), "gene"],
                      has_gene = ifelse(is.na(overlap_gene), FALSE, TRUE)) %>% group_by(has_gene) %>% summarise(n=n()) %>%
  ungroup() %>% mutate(prop = n/sum(n))

```


- do for width of gene??

```{r}
eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>% group_by(width) %>%
    mutate(num_bins_same_width = n()) %>%
    group_by(width, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_width) %>%
    distinct(width, num_bins_same_width, num_per_dmr, prop, has_dmr) %>%
    group_by(width) %>%
    .[order(.$width),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = width, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()

eh_genes %>% data.frame() %>% mutate(has_dmr = overlap_gene_dmr[match(.$gene, overlap_gene_dmr$gene), "gene"], 
                                     has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>% group_by(width) %>%
    mutate(num_bins_same_width = n()) %>%
    group_by(width, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_width) %>%
    distinct(width, num_bins_same_width, num_per_dmr, prop, has_dmr) %>%
    group_by(width) %>%
    .[order(.$width),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = log2(width), y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```

# more detailed gene features
```{r}
library(annotatr)
```

```{r}
annots_gene <- build_annotations(genome = 'hg19', annotations = c('hg19_basicgenes', 'hg19_genes_intergenic',
    'hg19_genes_intronexonboundaries'))
```

```{r}
annots_gene <- annots_gene %>% data.frame() %>% filter(seqnames %in% unique(bsseq_res$seqnames)) %>% as_granges()
```

```{r}
#anno_cpg_by_gene <- annotate_regions(as_granges(bsseq_res), annots_gene) %>% data.frame()
```

```{r}
bseq_dmrs_genes <- annotate_regions(as_granges(bsseq_dmrs), annots_gene) %>% data.frame()
```
```{r}

```

```{r}
bseq_dmrs_genes %>% group_by(annot.type) %>% summarise(n=n()) %>% ungroup() %>% mutate(prop = n/sum(n))
```
```{r}
bseq_dmrs_genes %>% distinct(rank, annot.type) %>% group_by(annot.type) %>% summarise(n_dmr = n(), prop_total_dmr = n()/69858)
```

```{r}
annots_gene %>% data.frame() %>% filter(type == "hg19_genes_promoters") %>% mutate(has_dmr = bseq_dmrs_genes[match(.$id, bseq_dmrs_genes$id), "rank"],
                                                                                         has_dmr = ifelse(!is.na(has_dmr), TRUE, FALSE)) %>%  
  
```
- oh wait all the width is the same


# cpg annotations - ISLANDS
```{r}
annotations %>% head()
```
```{r}
anno_islands <- annotations %>% data.frame() %>% filter(type == "hg19_cpg_islands") %>% mutate(island_num = 1:n())
island_cg <- find_overlaps(as_granges(anno_islands), as_granges(wgbs_counts)) %>% data.frame() %>% group_by(island_num) %>% summarise(num_cg = n()) %>% data.frame()
```
```{r}
head(anno_islands)
```

```{r}
find_overlaps(as_granges(eh_genes), as_granges(anno_islands)) %>% data.frame() %>% group_by(gene, num_cg.x) %>% summarise(n_islands = n()) %>%
  ggplot(aes(x = num_cg.x, y = n_islands)) +
  geom_point(alpha = 0.4)
```
```{r}
find_overlaps(as_granges(eh_genes), as_granges(anno_islands)) %>% data.frame() %>%
  group_by(width) %>% summarise(n_islands = n()) %>%
  ggplot(aes(x = log2(width), y = n_islands)) +
  geom_point(alpha = 0.4)
```


```{r}
anno_islands <- anno_islands %>% mutate(num_cg = island_cg[match(.$island_num, island_cg$island_num), "num_cg"])
anno_islands %>% group_by(num_cg) %>% summarise(n= n()) %>%
  ggplot(aes(x = num_cg, y = n)) +
  geom_point()
```

```{r}
bseq_dmrs_cpg_anno <- annotate_regions(as_granges(bsseq_dmrs), annotations) %>% data.frame()
```
```{r}
bseq_dmrs_cpg_isl <- filter(bseq_dmrs_cpg_anno, annot.type == "hg19_cpg_islands")
```

```{r}
anno_islands %>% mutate(has_dmr = ifelse(id %in% bseq_dmrs_cpg_isl$annot.id, TRUE, FALSE)) %>% group_by(num_cg) %>%
      mutate(num_bins_same_cg = n()) %>%
    group_by(num_cg, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_cg) %>%
    distinct(num_cg, num_bins_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg) %>%
    .[order(.$num_cg),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = num_cg, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```


- prop of dmrs that have island
```{r}
bsseq_dmrs %>% mutate(has_island = ifelse(rank %in% bseq_dmrs_cpg_isl$rank, TRUE, FALSE)) %>% group_by(has_island) %>% summarise(n=n()) %>% ungroup() %>% mutate(prop = n/sum(n))
```

#################
# repeat for biomart (has a tss)

```{r}
biomart_genes
bsseq_dmrs
```
- overlap genes and dmrs

```{r}
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame()
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% distinct(position, rank)
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% distinct(ensembl_gene_id, gene_name)
```
```{r}
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes), maxgap = 10000) %>% data.frame()
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes), maxgap = 10000) %>% data.frame() %>% distinct(position, rank)
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes), maxgap = 10000) %>% data.frame() %>% distinct(ensembl_gene_id, gene_name)
```
```{r}
pair_nearest(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% distinct(position, rank)

pair_nearest(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% group_by(position, rank) %>% summarise(n=n()) %>% .[order(.$n, decreasing = TRUE),]
```
- ok this gives 1-to-1 gene pairing and dmr
- now can get distance of dmr to tss
```{r}
pair_nearest(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame()
```
- but are we gene or dmr centric?
- probs gene because that's what we care about after
- ok so position of dmr relative to gene
- overlap we can get from find_overlaps
- need to take strand into account
- how did i do it in damsel?
- it was a minus thing and then a case_when
- but that dealing with 2 lots of ends, this is just tss

```{r}
pair_nearest(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% 
  mutate(tss_start = TSS - granges.x.start, tss_end = TSS - granges.x.end) %>% .[,c(1:3,10,24:26,28,34:35)] %>%
  mutate(tss_rel_peak = case_when(tss_start > 0 & tss_end > 0 & tss_end < tss_start ~ "Peak upstream", 
                                  tss_start < 0 & tss_end < 0 & tss_end < tss_start ~ "Peak downstream"),
         tss_rel_peak = case_when(tss_rel_peak == "Peak upstream" & granges.y.strand == "-" ~ "Peak downstream", 
                                  tss_rel_peak == "Peak downstream" & granges.y.strand == "-" ~ "Peak upstream",
                                  TRUE ~ tss_rel_peak))
pair_nearest(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% 
  mutate(tss_start = TSS - granges.x.start, tss_end = TSS - granges.x.end) %>% .[,c(1:3,10,24:26,28,34:35)] %>%
  mutate(tss_rel_peak = case_when(tss_start > 0 & tss_end > 0 & tss_end < tss_start ~ "Peak upstream", 
                                  tss_start < 0 & tss_end < 0 & tss_end < tss_start ~ "Peak downstream"),
         tss_rel_peak = case_when(tss_rel_peak == "Peak upstream" & granges.y.strand == "-" ~ "Peak downstream", 
                                  tss_rel_peak == "Peak downstream" & granges.y.strand == "-" ~ "Peak upstream",
                                  TRUE ~ tss_rel_peak)) %>% filter(is.na(tss_rel_peak))
```
```{r}
find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% filter(gene_name == "DOK3")
```
- ok the nas as suspected seem to match  the overlaps
```{r}
overlap_biomart_dmr <- find_overlaps(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% 
  mutate(combo = paste0(ensembl_gene_id, "-DMR", rank))
```

- i think i may have more na's than i think
- so more overlaps
- but pair_nearest does 1-to-1
- find overlaps does not
- so want to match the exact

```{r}
anno_biomart_dmr <- pair_nearest(as_granges(bsseq_dmrs), as_granges(biomart_genes)) %>% data.frame() %>% 
  mutate(tss_start = TSS - granges.x.start, tss_end = TSS - granges.x.end, combo = paste0(ensembl_gene_id, "-DMR", rank)) %>% mutate(tss_rel_peak = ifelse(combo %in% overlap_biomart_dmr$combo, "Overlap", NA),                                              tss_rel_peak = case_when(is.na(tss_rel_peak) & tss_start > 0 & tss_end > 0 &                                                                      tss_end < tss_start ~ "Peak_upstream",
                                  is.na(tss_rel_peak) & tss_start < 0 & tss_end < 0 & 
                                    tss_end < tss_start ~ "Peak_downstream", TRUE ~ tss_rel_peak),
         tss_rel_peak = case_when(tss_rel_peak == "Peak_upstream" & granges.y.strand == "-" ~ "Peak_downstream", 
                                  tss_rel_peak == "Peak_downstream" & granges.y.strand == "-" ~ "Peak_upstream",
                                  TRUE ~ tss_rel_peak),
         min_dist = case_when(tss_rel_peak == "Overlap" ~ 0, 
                              tss_rel_peak == "Peak_downstream" & granges.y.strand == "+" ~ tss_start,
                              tss_rel_peak == "Peak_downstream" & granges.y.strand == "-" ~ tss_end,
                              tss_rel_peak == "Peak_upstream" & granges.y.strand == "+" ~ tss_end,
                              tss_rel_peak == "Peak_upstream" & granges.y.strand == "-" ~ tss_start))
```
- ok so the overlap is to the gene body not the tss in particular but still

- ok now bin plot

```{r}
anno_biomart_dmr %>%
  ggplot(aes(x = min_dist)) +
  geom_density()
```
```{r}
filter(anno_biomart_dmr, abs(min_dist) < 5000)
```

```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(width) %>%
    mutate(num_bins_same_width = n()) %>%
    group_by(width, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_width) %>%
    distinct(width, num_bins_same_width, num_per_dmr, prop, has_dmr) %>%
    group_by(width) %>%
    .[order(.$width),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = width, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(width) %>%
    mutate(num_bins_same_width = n()) %>%
    group_by(width, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_width) %>%
    distinct(width, num_bins_same_width, num_per_dmr, prop, has_dmr) %>%
    group_by(width) %>%
    .[order(.$width),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = log2(width), y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```
- now in the way I've tried to fix it
- this is by width remember
- not num_cg

```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(width) %>%
    mutate(num_bins_same_width = n()) %>% 
  .[order(.$width),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(width_bin_group = sum(width)) %>%
  group_by(width_bin_group) %>%
    mutate(num_bins_group_same_width = n()) %>%
    group_by(width_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_width) %>%
    distinct(width_bin_group, num_bins_group_same_width, num_per_dmr, prop, has_dmr) %>%
    group_by(width_bin_group) %>%
    .[order(.$width_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = width_bin_group, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```
- well hello that's pretty

- now with num_cg

```{r}
biomart_cg <- find_overlaps(as_granges(biomart_genes), as_granges(wgbs_counts)) %>% data.frame() %>% group_by(ensembl_gene_id) %>% summarise(num_cg = n())
biomart_cg <- data.frame(biomart_cg)
```

```{r}
biomart_genes <- biomart_genes %>% mutate(num_cg = biomart_cg[match(.$ensembl_gene_id, biomart_cg$ensembl_gene_id), "num_cg"],
                                          num_cg = ifelse(is.na(num_cg), 0, num_cg))
anno_biomart_dmr <- anno_biomart_dmr %>% mutate(num_cg = biomart_cg[match(.$ensembl_gene_id, biomart_cg$ensembl_gene_id), "num_cg"], num_cg = ifelse(is.na(num_cg), 0, num_cg))
```
```{r}
anno_biomart_dmr$rank %in% bsseq_dmrs$rank %>% summary()
bsseq_dmrs$rank %in% filter(anno_biomart_dmr, tss_rel_peak == "Overlap")$rank %>% summary()
```
```{r}
49744/69858
```
```{r}
biomart_genes
```

```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>%
    group_by(num_cg, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_cg) %>%
    distinct(num_cg, num_bins_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg) %>%
    .[order(.$num_cg),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = num_cg, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>%
    group_by(num_cg, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_cg) %>%
    distinct(num_cg, num_bins_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg) %>%
    .[order(.$num_cg),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = log2(num_cg), y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```
```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>%
    group_by(num_cg, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_same_cg) %>%
    distinct(num_cg, num_bins_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg) %>%
    .[order(.$num_cg),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE)
```
- new version
```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>% 
  .[order(.$num_cg),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = sum(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = num_cg_bin_group, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) < 5000)$ensembl_gene_id, TRUE, FALSE)) %>%
  group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>% 
  .[order(.$num_cg),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = sum(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = num_cg_bin_group, y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```
- ok need to be able to explain what the code does
- in order to send off and find out if it's correct (or how wrong it is)
- we've said if a gene has a dmr or not - and defined that by overlap (or by min dist to dmr)
- we have num of cgs per gene
- order the genes by that
- group genes into bins of 100 genes
- got sum of cgs in that group (100 genes with 0 cgs - 0 cgs)
(side note - maybe sum is incorrect - maybe it should be something else??)
- then treated it as before
- so do any other bins have the sum number of cgs
- use this to get the proportion

```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>% 
  .[order(.$num_cg),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = sum(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = log2(num_cg_bin_group), y = prop)) +
      geom_point(alpha = 0.4) +
      geom_smooth()
```

- no we want average - not sum

```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>% 
  .[order(.$num_cg),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = mean(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = num_cg_bin_group, y = prop)) +
      geom_point(alpha = 0.4) 
  #    geom_smooth() 

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>% 
  .[order(.$num_cg),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = mean(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE) %>%
    ggplot(aes(x = log2(num_cg_bin_group), y = prop)) +
      geom_point(alpha = 0.4) 
 #     geom_smooth()
```
- alicia likes these plots - yay!
```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>% group_by(has_dmr) %>% summarise(n=n()) %>% ungroup() %>% mutate(ratio = n/sum(n))
```
```{r}
biomart_genes
```
```{r}
ann
```


```{r}
biomart_genes %>% mutate(n_genes = n()) %>% filter(num_cg == 0) %>% mutate(n_0 = n(), ratio = n()/n_genes)
```
- need to be able to explain the math well
```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  .[order(.$num_cg),] %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)])

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  .[order(.$num_cg),] %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = mean(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n())

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  .[order(.$num_cg),] %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = mean(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr)

biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(anno_biomart_dmr, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  .[order(.$num_cg),] %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = mean(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = n(), prop = n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE)
```

```{r}
biomart_genes %>% group_by(gene_name) %>% summarise(n=n()) %>% .[order(.$n, decreasing = TRUE),]
```
```{r}
biomart_cg
```

```{r}
biomart_cg_full <- find_overlaps(as_granges(biomart_genes), as_granges(wgbs_counts)) %>% data.frame()
```

```{r}
wgbs_counts[,1:4]
```
density = num_cg per gene/length of gene
-islands - too definition dependent
-ditch tss and biomart - just use start and be up 5k, down 2k
-check against belinda's code
-then start running goseq
