---
title: "03_genesToGST"
author: "Caitlin Page"
date: "2025-04-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Packages

```{r}
library(missMethyl)
library(goseq)
library(org.Hs.eg.db)

library(dplyr)
library(ggplot2)
```

# copied code over from 00_exploreCpGGST.Rmd 
want focus there to be on exploring, here on the problem of converting ensembl to entrez etc
```{r}
x <- org.Hs.egGO
# Get the entrez gene identifiers that are mapped to a GO ID
entrez_map <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[entrez_map])

xx <- as.list(org.Hs.egGO2ALLEGS)

go_entrez <- data.frame(cbind(go = gsub("\\..*", "", names(unlist(xx))), entrez = unlist(xx)))

# map entrez to gene symbol https://web.mit.edu/~r/current/arch/i386_linux26/lib/R/library/org.Hs.eg.db/html/org.Hs.egSYMBOL.html
x <- org.Hs.egSYMBOL
# Get the gene symbol that are mapped to an entrez gene identifiers
mapped_genes <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[mapped_genes])

entrez_symbol <- data.frame(cbind(entrez = gsub("\\..*", "", names(unlist(xx))), symbol = unlist(xx)))
go_entrez <- go_entrez %>% mutate(symbol = entrez_symbol[match(.$entrez, entrez_symbol$entrez), "symbol"])
go_entrez <- go_entrez %>% mutate(ontology = go_info[match(.$go, go_info$GOID), "ONTOLOGY"],
                                  term = go_info[match(.$go, go_info$GOID), "TERM"])
go_entrez %>% filter(entrez == 1)
```
* theoretically can get positions for org package but I think it's icky
* and no idea what genome it would follow
* do a thing for txdb
* but right now just carry over biomart
https://support.bioconductor.org/p/p133188/
```{r}
biomart_genes <- biomart_genes %>% mutate(coords = paste0(seqnames, ":", start, "-", end))
```
```{r}
go_entrez <- go_entrez %>% mutate(coords = biomart_genes[match(.$symbol, biomart_genes$gene_name), "coords"],
                                  seqnames = biomart_genes[match(.$symbol, biomart_genes$gene_name), "seqnames"],
                                  start = biomart_genes[match(.$symbol, biomart_genes$gene_name), "start"],
                                  end = biomart_genes[match(.$symbol, biomart_genes$gene_name), "end"],
                                  n_cg_gene = biomart_genes[match(.$symbol, biomart_genes$gene_name), "num_cg"]) 
go_entrez %>% distinct(entrez, symbol, coords) %>% filter(is.na(coords))
```

* quick side note - how much does this impact??
```{r}
biomart_genes %>% filter(gene_name == "twnk")
biomart_genes %>% filter(ensembl_gene_id == "ENSG00000107815")
```
* same gene - different names
```{r}
go_entrez %>% filter(symbol == "C10orf2")
```

* this could be the case for all of them
 https://support.bioconductor.org/p/9148989/
* no clear solution
* as basically everyone hates this kind of stuff - read a recc which was don't switch just stick to what works for you
* maybe mapIds - different function
```{r}
go_entrez <- go_entrez %>% group_by(go) %>% mutate(n_genes_go = n()) %>% ungroup()
```
```{r}
go_entrez_filt <- go_entrez %>% filter(!is.na(start)) %>% group_by(go) %>% mutate(n_genes_go = n(), av_cg_go = mean(n_cg_gene), sum_cg_go = sum(n_cg_gene), gene_prop = sum(n_cg_gene)/n()) %>% ungroup()

go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go) %>%
  ggplot(aes(x = n_genes_go, y = av_cg_go)) +
  geom_point(alpha = 0.3)
```

```{r}
go_entrez_filt %>% distinct(go, term, n_genes_go, sum_cg_go) %>%
  ggplot(aes(x = n_genes_go, y = sum_cg_go)) +
  geom_point(alpha = 0.3)

go_entrez_filt %>% distinct(go, term, n_genes_go, gene_prop) %>%
  ggplot(aes(x = n_genes_go, y = gene_prop)) +
  geom_point(alpha = 0.3)
```
```{r}
go_entrez_filt_simp <- go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go, sum_cg_go, gene_prop)
go_entrez_filt_simp %>% filter(av_cg_go == 0)
```

```{r}
cor <- cor.test(go_entrez_filt_simp$n_genes_go, go_entrez_filt_simp$sum_cg_go)
cor$statistic
```


```{r}
go_entrez_filt_simp %>% group_by(n_genes_go) %>% summarise(n_cat = n())
```

```{r}
go_entrez_filt_simp %>% group_by(n_genes_go) %>% mutate(n_cat = n()) %>%
  ggplot(aes(x = n_genes_go, y = n_cat)) +
  geom_point() +
  scale_x_continuous(limits = c(0,500))

go_entrez_filt_simp %>% group_by(n_genes_go) %>% mutate(n_cat = n()) %>%
  ggplot(aes(x = n_genes_go, y = log2(n_cat))) +
  geom_point() +
  scale_x_continuous(limits = c(0,500))
```
* more categories with not many genes
```{r}
go_entrez_filt_simp %>% 
  ggplot(aes(x = n_genes_go, y = av_cg_go)) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(limits = c(0,500))

go_entrez_filt_simp %>% 
  ggplot(aes(x = n_genes_go, y = av_cg_go)) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(limits = c(0,250))

go_entrez_filt_simp %>% 
  ggplot(aes(x = n_genes_go, y = log2(av_cg_go))) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(limits = c(0,500))

go_entrez_filt_simp %>% 
  ggplot(aes(x = n_genes_go, y = log2(av_cg_go))) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(limits = c(0,250))
```
```{r}
go_entrez_filt_simp %>%
  ggplot(aes(x = sum_cg_go, y = av_cg_go)) +
  geom_point()
```
```{r}
go_entrez_filt_simp %>% .[order(.$av_cg_go, decreasing = TRUE),]
go_entrez_filt_simp %>% .[order(.$sum_cg_go, decreasing = TRUE),]
go_entrez_filt_simp %>% .[order(.$gene_prop, decreasing = TRUE),]
go_entrez_filt_simp %>% .[order(.$n_genes_go, decreasing = TRUE),]
```
* not suprised terms with most genes, and sum cg are the same - and very vague
* the top gene prop and av cg are also the same - and these are just those weird genes that have heaps of cpgs in them - as they also have only 1 or 2 genes in them

```{r}
go_entrez_filt_simp %>%
  ggplot(aes(x = log2(n_genes_go))) +
  geom_density()

go_entrez_filt_simp %>%
  ggplot(aes(x = av_cg_go)) +
  geom_density()

go_entrez_filt_simp %>%
  ggplot(aes(x = log2(av_cg_go))) +
  geom_density()

go_entrez_filt_simp %>%
  ggplot(aes(x = sum_cg_go)) +
  geom_density()

go_entrez_filt_simp %>%
  ggplot(aes(x = log2(sum_cg_go))) +
  geom_density()

go_entrez_filt_simp %>%
  ggplot(aes(x = gene_prop)) +
  geom_density()

go_entrez_filt_simp %>%
  ggplot(aes(x = log2(gene_prop))) +
  geom_density()
```

```{r}
go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go)
```

```{r}
go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go) %>% mutate(min(n_genes_go), max(n_genes_go), mean(n_genes_go), sd(n_genes_go))

go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go) %>% group_by(n_genes_go) %>% mutate(n_same=n(), av_same = mean(av_cg_go)) %>%
  ggplot(aes(x = n_same, y = av_same)) +
  geom_point(alpha = 0.3)
```

```{r}
go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go) %>% .[order(.$n_genes_go),] %>% 
  ggplot(aes(x = as.factor(n_genes_go), y = av_cg_go)) +
  geom_boxplot()
```


```{r}
go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go) %>% filter(n_genes_go <= 10000) %>%
  ggplot(aes(x = log10(n_genes_go), y = log10(av_cg_go))) +
  geom_point(alpha = 0.3)
```


```{r}
go_entrez_filt %>% distinct(go, term, n_genes_go, av_cg_go)
```



```{r}
overlap_cpg <- find_overlaps(as_granges(biomart_genes), as_granges(wgbs_counts[,1:4])) %>% data.frame()
overlap_cpg %>% group_by(ensembl_gene_id, gene_name, num_cg) %>% summarise(n= n()) %>%
  ggplot(aes(x = num_cg, y = n)) +
  geom_point()
```

```{r}
biomart_genes
distinct(biomart_genes, ensembl_gene_id)
distinct(biomart_genes, gene_name)
```


- have 3,151 genes with no coords - not good - but right now live with it - how many 

```{r}
x <- org.Hs.egCHRLOC
# Get the gene symbol that are mapped to an entrez gene identifiers
mapped_genes <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[mapped_genes])

data.frame(cbind(entrez = gsub("\\..*", "", names(unlist(xx))), symbol = unlist(xx)))

x <- org.Hs.egCHR
# Get the gene symbol that are mapped to an entrez gene identifiers
mapped_genes <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[mapped_genes])

data.frame(cbind(entrez = gsub("\\..*", "", names(unlist(xx))), symbol = unlist(xx)))

x <- org.Hs.egCHRLOCEND
# Get the gene symbol that are mapped to an entrez gene identifiers
mapped_genes <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[mapped_genes])

data.frame(cbind(entrez = gsub("\\..*", "", names(unlist(xx))), symbol = unlist(xx)))
```
```{r}
biomart_genes %>% filter(gene_name == "A1BG")
```

```{r}
org.Hs.eg
```

```{r}
.getGO <- function(){
  if(!requireNamespace("org.Hs.eg.db", quietly = TRUE))
    stop("org.Hs.eg.db package required but not installed.")
  egGO2ALLEGS <- utils::getFromNamespace("org.Hs.egGO2ALLEGS", "org.Hs.eg.db")
  GeneID.PathID <- AnnotationDbi::toTable(egGO2ALLEGS)[,c("gene_id", "go_id", "Ontology")]
  d <- !duplicated(GeneID.PathID[, c("gene_id", "go_id")])
  GeneID.PathID <- GeneID.PathID[d, ]
  GOID.TERM <- suppressMessages(AnnotationDbi::select(GO.db::GO.db,
                                                      keys=unique(GeneID.PathID$go_id),
                                                      columns=c("GOID","ONTOLOGY","TERM"),
                                                      keytype="GOID"))
  go <- tapply(GeneID.PathID$gene_id, GeneID.PathID$go_id, list)

  list(idList=go, idTable=GOID.TERM)
}
```

* ok so go_entrez is used in the comparing
* and get_go is used in the testing
* both do the go2allegs
* which allows mapping of gene to term, or to any of its child terms
* i think I have an issue with the allowing mapping to any of the child terms
```{r}
org.Hs.egGO
```

```{r}

```

* the version of getting the go terms
