---
title: "04_compareStatTests"
author: "Caitlin Page"
date: "2025-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Packages

```{r}
library(BiasedUrn)
library(missMethyl)
library(goseq)
library(GO.db)

library(rbenchmark)

library(dplyr)
library(ggVennDiagram)
library(ggplot2)
```

## Data
* need inputs for GST
```{r}
DMPs <- readRDS("../output/array_dmps.rds")
```


## Microarray

```{r}
topCpGs <- DMPs[1:10000,]
sigCpGs <- rownames(topCpGs)
```
```{r}
mapped_entrez <- getMappedEntrezIDs(sig.cpg=sigCpGs, all.cpg=rownames(DMPs))
```

```{r}
gst <- gometh(sig.cpg=sigCpGs, all.cpg=rownames(DMPs), collection="GO",
              plot.bias=TRUE)
```

```{r}
sig_terms_array <- topGSA(gst, n = 1000) %>% filter(FDR < 0.05)
sig_terms_array
sig_terms_array <- readRDS("../output/mm_array_res.rds")
```


# Goseq version to compare approaches
- except I need the genes and that's under the hood
```{r}
mapped_entrez$freq %>% data.frame()
```
```{r}
length(mapped_entrez$sig.eg)
length(mapped_entrez$de)
length(mapped_entrez$universe)
```

```{r}
gene.vector <- mapped_entrez$de
names(gene.vector) <- mapped_entrez$universe

pwf <- goseq::nullp(gene.vector, "hg19", "Entrez", bias.data = data.frame(mapped_entrez$freq)$Freq)
pwf
GO.wall <- goseq::goseq(pwf, "hg19", "Entrez")
GO.wall <- GO.wall %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))
GO.wall
```
```{r}
genes_symbol <- AnnotationDbi::select(org.Hs.eg.db, 
                                keys=AnnotationDbi::keys(org.Hs.eg.db), 
                                columns=c("ENTREZID","SYMBOL"), 
                                keytype="ENTREZID")

```

```{r}
universe <- data.frame(mapped_entrez$freq)
universe$de <- mapped_entrez$de
universe$eg.all <- as.character(universe$eg.all)
universe$symbol <- genes_symbol[match(universe$eg.all, genes_symbol$ENTREZID), "SYMBOL"]
universe
```


```{r}
gene.vector <- universe$de
names(gene.vector) <- universe$symbol
pwf <- goseq::nullp(gene.vector, "hg19", "geneSymbol", bias.data = universe$Freq)
pwf
GO.wall <- goseq::goseq(pwf, "hg19", "geneSymbol")
GO.wall <- GO.wall %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))
GO.wall
```
```{r}
GO.samp <- goseq(pwf,"hg19","geneSymbol",method="Sampling",repcnt=1000)
```
```{r}
GO.samp <- GO.samp %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))
```

```{r}
GO.fish <- goseq_mod(pwf,"hg19","geneSymbol",method="Fishers")
GO.fish <- GO.fish %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))
```
```{r}
system.time(goseq(pwf, "hg19", "geneSymbol"))
system.time(goseq(pwf,"hg19","geneSymbol",method="Sampling",repcnt=1000))
system.time(goseq_mod(pwf,"hg19","geneSymbol",method="Fishers"))
```
```{r}
benchmark(wall = goseq(pwf, "hg19", "geneSymbol"), fish = goseq_mod(pwf,"hg19","geneSymbol",method="Fishers"), replications = 50)
```
* interesting - basically no difference in timing

## compare results - fishers and wallenius
* can I do a correlation?
* because that's basically just comparing how similar 2 lists are
* if wants numbers - can just give diff terms diff numbers
* ok its more an association thing but it could still work

```{r}
GO.wall <- GO.wall %>% .[order(.$FDR),] %>% mutate(wall_rank = 1:n())
GO.samp <- GO.samp %>% .[order(.$FDR),] %>% mutate(samp_rank = 1:n())
GO.fish <- GO.fish %>% .[order(.$FDR),] %>% mutate(fish_rank = 1:n())
```

```{r}
GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = samp)) +
  geom_point()

GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish)) +
  geom_point()

GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = samp, y = fish)) +
  geom_point()
```
```{r}
ggVennDiagram(list(wall = filter(GO.wall, FDR < 0.05)$category, 
              samp = filter(GO.samp, FDR < 0.05)$category,
              fish = filter(GO.fish, FDR < 0.05)$category))
```


```{r}
GO.wall 
GO.samp
GO.fish
```
* GO.samp - the results after adjustment are not significant
* GOseq PDF: "However,
the accuracy of the sampling method is limited by the number of samples generated, repcnt, such
that very low p-values will not be correctly calculated. Significantly enriched GO terms may then
be missed after correcting for multiple testing."
* this may be why nothing is significant
* but wasn't the sampling supposed to be the best version - and Wallenius an approximation?
```{r}
plot(log10(GO.wall[,2]), log10(GO.samp[match(GO.wall[,1],GO.samp[,1]),2]),
xlab="log10(Wallenius p-values)",ylab="log10(Sampling p-values)",
xlim=c(-3,0))
abline(0,1,col=3,lty=2)
```

## WGBS
```{r}
wgbs_counts <- readRDS("../output/wgbs_counts.rds")
bsseq_dmrs <- readRDS("../output/bsseq_dmrs.rds")
biomart_genes <- readRDS("../output/genes_biomart.rds")
```
```{r anno}
source("../code/function_binBiasPlot.R")
test_anno <- annoGeneDmr(wgbs_counts, bsseq_dmrs, gene_source = "biomaRt", genes = biomart_genes)
```
```{r}
test_anno_filt <- test_anno %>% filter(areaStat > 100)
```


```{r}
test_bias <- biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(test_anno, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
    group_by(num_cg) %>%
    mutate(num_bins_same_cg = dplyr::n()) %>%
    .[order(.$num_cg),] %>%
    ungroup() %>%
    mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>%
    group_by(bin_group) %>%
    mutate(num_cg_bin_group = mean(num_cg)) %>%
    ungroup() %>% mutate(has_dmr = ifelse(has_dmr == FALSE, 0, 1))
test_bias
test_bias_filt <- biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(test_anno_filt, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
    group_by(num_cg) %>%
    mutate(num_bins_same_cg = dplyr::n()) %>%
    .[order(.$num_cg),] %>%
    ungroup() %>%
    mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>%
    group_by(bin_group) %>%
    mutate(num_cg_bin_group = mean(num_cg)) %>%
    ungroup() %>% mutate(has_dmr = ifelse(has_dmr == FALSE, 0, 1))
test_bias_filt
```

# goseq
```{r}
gene.vector <- test_bias_filt$has_dmr
names(gene.vector) <- test_bias_filt$ensembl_gene_id

pwf_seq <- goseq::nullp(gene.vector, "hg19", "ensGene", bias.data = test_bias_filt$num_cg_bin_group)
pwf_seq
GO.wall_seq <- goseq::goseq(pwf_seq, "hg19", "ensGene")
GO.wall_seq <- GO.wall_seq %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))

GO.samp_seq <- goseq(pwf_seq,"hg19","ensGene",method="Sampling",repcnt=1000)
GO.samp_seq <- GO.samp_seq %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))

GO.fish_seq <- goseq_mod(pwf_seq,"hg19","ensGene",method="Fishers")
GO.fish_seq <- GO.fish_seq %>% dplyr::mutate(FDR = stats::p.adjust(.data$over_represented_pvalue, method = "BH"))

GO.wall_seq <- GO.wall_seq %>% .[order(.$FDR),] %>% mutate(wall_rank = 1:dplyr::n())
GO.samp_seq <- GO.samp_seq %>% .[order(.$FDR),] %>% mutate(samp_rank = 1:dplyr::n())
GO.fish_seq <- GO.fish_seq %>% .[order(.$FDR),] %>% mutate(fish_rank = 1:dplyr::n())
```

```{r}
system.time(goseq(pwf_seq, "hg19", "ensGene"))
system.time(goseq(pwf_seq,"hg19","ensGene",method="Sampling",repcnt=1000))
system.time(goseq_mod(pwf_seq,"hg19","ensGene",method="Fishers"))
```

```{r}
benchmark(wall = goseq(pwf_seq, "hg19", "ensGene"), fish = goseq_mod(pwf_seq,"hg19","ensGene",method="Fishers"), replications = 50)
```


# missMethyl
```{r}
test_out <- readRDS("../output/mm_fn_out_num.rds")
test_out_filt <- run_miss_methyl_a(biomart_genes, filter(test_anno, abs(areaStat) > 100))
```

```{r}
MM_fish <- run_miss_methyl_b(test_out, method = "Fishers")
```

```{r}
system.time(run_miss_methyl_b(test_out, method = "Wallenius"))
system.time(run_miss_methyl_b(test_out, method = "Fishers"))
```
* ok this implies no difference still
```{r}
system.time(run_miss_methyl_b(test_out_filt, method = "Wallenius"))
system.time(run_miss_methyl_b(test_out_filt, method = "Fishers"))
```
* damn now it's not even faster?
* trying to skip testing for 0 de for wallenius only
```{r}
system.time(run_miss_methyl_b(test_out_filt, method = "Wallenius"))
system.time(run_miss_methyl_b(test_out_filt, method = "Fishers"))
```

* damn that made it longer than the previous - and worse it made the system time longer
* ok but the other one took longer this time as well

* but also thought of better way - just remove them before the testing
```{r}
system.time(run_miss_methyl_b(test_out_filt, method = "Wallenius"))
system.time(run_miss_methyl_b(test_out_filt, method = "Fishers"))
```
* haha that is faster system wise
* still slow overall
* still don't get why it's so much slower than goseq
* and now fishers is slower than wall again
* why is it so much slower than goseq??
```{r}
GO.wall
```

```{r}
run_miss_methyl_b(test_out_filt, method = "Wallenius")
```

```{r}
test_out$freq %>% data.frame()
test_out$de %>% data.frame()
```
```{r}
universe
```
```{r}
GO.fish %>% filter(numDEInCat == 0)
```

```{r}
MM_fish %>% filter(DE == 0) %>% .[order(.$N),]
```



## Code
# seq code - missmethyl

* if Fishers is independent, I should be able to not do it as a loop or whatever
* so I can have a vector for x - which is the num of red balls sampled
* but we put in the de? - right the number of de genes
* but this wouldn't change? because it's our list? - or its got something to do with something
* maybe it still needs to be a loop because it's still an individual test for each category and we need a p val for each
* and the speed should be that something else is happening under the hood of the test which might be faster
* or maybe I can // it?
*I'm not figuring out what goes under the hood of the tests - found the code - I think it's written in C++
* I do not read C++
* can say the Wallenius file is longer than the Fishers
* ok so reviewed the what goes in and out and that all makes sense
* guessing minimal speed difference in methods is because of the c++ efficiency

# what to do for speed?
* prefilter out go categories?? - so not test them if low?
* can't tell if that's what is currently happening in goseq - all 1 for p, mm still has a p


  if(prior.prob){
    for(i in 1:length(collection)){
      InSet <- eg.universe %in% collection[[i]]
      pw.red <- sum(pwf[InSet])/results[i,"N"]
      pw.white <- sum(pwf[!InSet])/(Nuniverse-results[i,"N"])
      odds <- pw.red/pw.white
      results[i,"P.DE"] <- BiasedUrn::pWNCHypergeo(results[i,"DE"],
                                                   results[i,"N"],
                                                   Nuniverse-results[i,"N"],
                                                   m,odds,lower.tail=FALSE) + 
          BiasedUrn::dWNCHypergeo(results[i,"DE"],
                                  results[i,"N"],
                                  Nuniverse-results[i,"N"],
                                  m,odds)
  }
  }
```{r}
system.time(go_stuff(test_out_filt))
```
```{r}
rbind(MM_fish[MM_fish$DE != 0,], MM_fish[MM_fish$DE == 0,])
```
```{r}
try_res <- try_speed(test_out_filt)
```
```{r}
class(try_res)
```

```{r}
try_res %>% data.frame() %>% .[.$DE != 0,] %>% as.matrix() %>% class()
data.frame(NA,ncol=4,nrow=10)
```
```{r}
try_collect <- try_speed(test_out_filt)
```

```{r}
go_stuff <- function(out) {
  sorted.eg.sig <- out$sig.eg #entrez ids for sig cpgs - genes that have sig cpgs
  eg.universe <- out$universe # entrez ids all cpgs(?) - but it's just all genes
  freq_genes <- out$freq
  test.de <- out$de #res for each gene
  #frac <- out$fract.counts
  equiv <- out$equiv
  ## setting up stuff
  # Check collection is a list with character vectors
  go <- .getGO()
  collection <- go$idList  
  #only if it's not a list
  #collection <- list(collection=go$idList)
  collection <- lapply(collection, as.character)
  # Make sure gene set collections don't have any NAs
  collection <- lapply(collection, function(x) x[!is.na(x)])
  # Remove genes that are NOT in the universe from collections
  collection <- lapply(collection, function(x) x[x %in% eg.universe])
  # Remove collections with no genes left after universe filter
  inUniv <- sapply(collection, function(x) length(x) > 0)
  collection <- collection[inUniv]
}

try_speed <- function(out) {
  sorted.eg.sig <- out$sig.eg #entrez ids for sig cpgs - genes that have sig cpgs
  eg.universe <- out$universe # entrez ids all cpgs(?) - but it's just all genes
  freq_genes <- out$freq
  test.de <- out$de #res for each gene
  #frac <- out$fract.counts
  equiv <- out$equiv
  ## setting up stuff
  # Check collection is a list with character vectors
  go <- .getGO()
  collection <- go$idList  
  #only if it's not a list
  #collection <- list(collection=go$idList)
  collection <- lapply(collection, as.character)
  # Make sure gene set collections don't have any NAs
  collection <- lapply(collection, function(x) x[!is.na(x)])
  # Remove genes that are NOT in the universe from collections
  collection <- lapply(collection, function(x) x[x %in% eg.universe])
  # Remove collections with no genes left after universe filter
  inUniv <- sapply(collection, function(x) length(x) > 0)
  collection <- collection[inUniv]
  
  
  results <- matrix(NA,ncol=4,nrow=length(collection))
  colnames(results) <- c("N", "DE", "P.DE","FDR")
  rownames(results) <- names(collection)
  results[,"N"] <- unlist(lapply(collection,length))
  SigGenesInSet <- rep(NA,length(collection)) # this one is optional
  
  results[,"DE"] <- unlist(lapply(collection, function(x) 
          sum((sorted.eg.sig %in% x))))
  Nuniverse <- length(eg.universe)
  m <- length(sorted.eg.sig)
  collection
}
```
```{r}
data.frame(try_res) %>% .[.$DE != 0,] %>% as.matrix() %>% nrow()
```
```{r}
length(try_collect)
length(try_collect[data.frame(try_res) %>% .[.$DE != 0,] %>% rownames()])
```

```{r}
.getGO <- function(){
  if(!requireNamespace("org.Hs.eg.db", quietly = TRUE))
    stop("org.Hs.eg.db package required but not installed.")
  egGO2ALLEGS <- utils::getFromNamespace("org.Hs.egGO2ALLEGS", "org.Hs.eg.db")
  GeneID.PathID <- AnnotationDbi::toTable(egGO2ALLEGS)[,c("gene_id", "go_id", "Ontology")]
  d <- !duplicated(GeneID.PathID[, c("gene_id", "go_id")])
  GeneID.PathID <- GeneID.PathID[d, ]
  GOID.TERM <- suppressMessages(AnnotationDbi::select(GO.db::GO.db, 
                                                      keys=unique(GeneID.PathID$go_id), 
                                                      columns=c("GOID","ONTOLOGY","TERM"), 
                                                      keytype="GOID"))
  go <- tapply(GeneID.PathID$gene_id, GeneID.PathID$go_id, list)
    
  list(idList=go, idTable=GOID.TERM)
}

.estimatePWF <- function(D,bias)
  # An alternative to goseq function nullp, which is transformation invariant
  # Belinda Phipson and Gordon Smyth
  # 6 March 2015
{
  prior.prob <- bias
  o <- order(bias)
  prior.prob[o] <- limma::tricubeMovingAverage(D[o],span=0.5)
  prior.prob
}

run_miss_methyl_b <- function(out, method, plot_bias=FALSE, ...) {
  sorted.eg.sig <- out$sig.eg #entrez ids for sig cpgs - genes that have sig cpgs
  eg.universe <- out$universe # entrez ids all cpgs(?) - but it's just all genes
  freq_genes <- out$freq
  test.de <- out$de #res for each gene
  #frac <- out$fract.counts
  equiv <- out$equiv
  ## setting up stuff
  # Check collection is a list with character vectors
  go <- .getGO()
  collection <- go$idList  
  #only if it's not a list
  #collection <- list(collection=go$idList)
  collection <- lapply(collection, as.character)
  # Make sure gene set collections don't have any NAs
  collection <- lapply(collection, function(x) x[!is.na(x)])
  # Remove genes that are NOT in the universe from collections
  collection <- lapply(collection, function(x) x[x %in% eg.universe])
  # Remove collections with no genes left after universe filter
  inUniv <- sapply(collection, function(x) length(x) > 0)
  collection <- collection[inUniv]
  
  
  results <- matrix(NA,ncol=4,nrow=length(collection))
  colnames(results) <- c("N", "DE", "P.DE","FDR")
  rownames(results) <- names(collection)
  results[,"N"] <- unlist(lapply(collection,length))
  SigGenesInSet <- rep(NA,length(collection)) # this one is optional
  
  results[,"DE"] <- unlist(lapply(collection, function(x) 
          sum((sorted.eg.sig %in% x))))
  Nuniverse <- length(eg.universe)
  m <- length(sorted.eg.sig)
  
  results <- data.frame(results)
  results_null <- results[results$DE == 0,]
  results_null$P.DE <- 1
  results_null$FDR <- 1
  
  results <- results[results$DE != 0,]
  collection <- collection[rownames(results)]
  results <- as.matrix(results)
  
  if(method != "None") {
  test_pwf <- .estimatePWF(D=test.de,bias=as.vector(freq_genes))
  #  if(plot_bias == TRUE) {
   #   plotBiasGrouped(genes = genes, anno_genes = , ...)
  #  }
  # the testing
  }
  if(method == "Wallenius") {
  results[, c("P.DE")] <- 
  t(sapply(1:length(collection), function(i) {
    # try to skip the 0 des to speed it up
 #   if(results[i,"DE"] == 0) {
 #     p_val <- 1
  #  } else {
    InSet <- eg.universe %in% collection[[i]]
    pw.red <- sum(test_pwf[InSet], na.rm = TRUE) / results[i, "N"] # so i don't know why I get na's - or do i?
    pw.white <- sum(test_pwf[!InSet], na.rm = TRUE) / (Nuniverse - results[i, "N"]) #anyway the na removal is not in original code
    odds <- pw.red / pw.white
    p_val <- BiasedUrn::pWNCHypergeo(results[i,"DE"],
                                                   results[i,"N"],
                                                   Nuniverse-results[i,"N"],
                                                   m,odds,lower.tail=FALSE) + 
          BiasedUrn::dWNCHypergeo(results[i,"DE"],
                                  results[i,"N"],
                                  Nuniverse-results[i,"N"],
                                  m,odds)
    #the optional thing
      # Get gene symbols of significant genes
    SigGenesEntrezID <- sorted.eg.sig[sorted.eg.sig %in% collection[[i]]]
    SigGenesSymbol <- suppressMessages(AnnotationDbi::select(org.Hs.eg.db, 
                                             keys = SigGenesEntrezID,
                                             columns = "SYMBOL"))
    SigGenesInSet[i] <- paste(SigGenesSymbol$SYMBOL,collapse=",")
    if(p_val==0){
        message("Pvalue of exactly zero detected. Performing hypergeometric 
                test for gene set ", rownames(results)[i])
        p_val <- stats::phyper(q=results[i,"DE"]-0.5,m=m,
                                           n=Nuniverse-m,k=results[i,"N"],
                                           lower.tail=FALSE)
    }
  #  }
    c(p_val)  # return the values as a vector
    })) #chatgpt translated loop to apply
  } else if(method == "None") {
    results[, c("P.DE")] <- 
    t(sapply(1:length(collection), function(i) {
      p_val <- stats::phyper(q=results[i,"DE"]-0.5,m=m,
                                           n=Nuniverse-m,k=results[i,"N"],
                                           lower.tail=FALSE)
      #the optional thing
      # Get gene symbols of significant genes
    SigGenesEntrezID <- sorted.eg.sig[sorted.eg.sig %in% collection[[i]]]
    SigGenesSymbol <- suppressMessages(AnnotationDbi::select(org.Hs.eg.db, 
                                             keys = SigGenesEntrezID,
                                             columns = "SYMBOL"))
    SigGenesInSet[i] <- paste(SigGenesSymbol$SYMBOL,collapse=",")
      c(p_val)
    }))
  } else if(method == "Fishers") {
    results[, c("P.DE")] <- 
  t(sapply(1:length(collection), function(i) {
    InSet <- eg.universe %in% collection[[i]]
    pw.red <- sum(test_pwf[InSet], na.rm = TRUE) / results[i, "N"] # so i don't know why I get na's - or do i?
    pw.white <- sum(test_pwf[!InSet], na.rm = TRUE) / (Nuniverse - results[i, "N"]) #anyway the na removal is not in original code
    odds <- pw.red / pw.white
    p_val <- BiasedUrn::pFNCHypergeo(results[i,"DE"],
                                                   results[i,"N"],
                                                   Nuniverse-results[i,"N"],
                                                   m,odds,lower.tail=FALSE) + 
          BiasedUrn::dFNCHypergeo(results[i,"DE"],
                                  results[i,"N"],
                                  Nuniverse-results[i,"N"],
                                  m,odds)
    #the optional thing
      # Get gene symbols of significant genes
    SigGenesEntrezID <- sorted.eg.sig[sorted.eg.sig %in% collection[[i]]]
    SigGenesSymbol <- suppressMessages(AnnotationDbi::select(org.Hs.eg.db, 
                                             keys = SigGenesEntrezID,
                                             columns = "SYMBOL"))
    SigGenesInSet[i] <- paste(SigGenesSymbol$SYMBOL,collapse=",")
    if(p_val==0){
        message("Pvalue of exactly zero detected. Performing hypergeometric 
                test for gene set ", rownames(results)[i])
        p_val <- stats::phyper(q=results[i,"DE"]-0.5,m=m,
                                           n=Nuniverse-m,k=results[i,"N"],
                                           lower.tail=FALSE)
    }
    c(p_val)  # return the values as a vector
    }))
  }
  
  
  results <- results %>% data.frame()
  results[,"FDR"] <- stats::p.adjust(results[,"P.DE"],method="BH")
  results[,"DE"] <- floor(results[,"DE"])
  results <- rbind(results, results_null)
  results <- data.frame(results, SigGenesInSet)
  
  results <- merge(go$idTable,results,by.x="GOID",by.y="row.names")
  rownames(results) <- results$GOID
  results <- results %>% .[order(.$FDR),]
  results
  
}

```

# microarray code
# goseq code
```{r}
reversemapping=function(map){
	tmp=unlist(map,use.names=FALSE)
	names(tmp)=rep(names(map),times=as.numeric(summary(map)[,1]))
	return(split(names(tmp),as.vector(tmp)))
}
goseq_mod <- function(pwf,genome,id,gene2cat=NULL,test.cats=c("GO:CC","GO:BP","GO:MF"),method="Wallenius",repcnt=2000,use_genes_without_cat=FALSE){
	################# Input pre-processing and validation ###################
	#Do some validation of input variables
	if(any(!test.cats%in%c("GO:CC","GO:BP","GO:MF","KEGG"))){
		stop("Invalid category specified.  Valid categories are GO:CC, GO:BP, GO:MF or KEGG")
	}
	if((missing(genome) | missing(id))){
		if(is.null(gene2cat)){
			stop("You must specify the genome and gene ID format when automatically fetching gene to GO category mappings.")
		}
		#If we're using user specified mappings, this obviously isn't a problem
		genome='dummy'
		id='dummy'
	}
	if(!any(method%in%c("Wallenius","Sampling","Hypergeometric", "Fishers"))){
		stop("Invalid calculation method selected.  Valid options are Wallenius, Sampling & Hypergeometric.")
	}
	if(!is.null(gene2cat) && (!is.data.frame(gene2cat) & !is.list(gene2cat))){
		stop("Was expecting a dataframe or a list mapping categories to genes.  Check gene2cat input and try again.")
	}
		
	#Factors are evil
	pwf=unfactor(pwf)
	gene2cat=unfactor(gene2cat)

	###################### Data fetching and processing ########################
	if(is.null(gene2cat)){
		#When we fetch the data using getgo it will be in the list format
		message("Fetching GO annotations...")
		gene2cat=getgo(rownames(pwf),genome,id,fetch.cats=test.cats)
		names(gene2cat)=rownames(pwf)
		#Do the two rebuilds to remove any nulls
		cat2gene=reversemapping(gene2cat)
		gene2cat=reversemapping(cat2gene)
	}else{
		#The gene2cat input accepts a number of formats, we need to check each of them in term
		message("Using manually entered categories.")
		#The options are a flat mapping (that is a data frame or matrix) or a list, where the list can be either gene->categories or category->genes
		if(class(gene2cat)!="list"){
			#it's not a list so it must be a data.frame, work out which column contains the genes
			genecol_sum=as.numeric(apply(gene2cat,2,function(u){sum(u%in%rownames(pwf))}))
			genecol=which(genecol_sum!=0)
			if(length(genecol)>1){
				genecol=genecol[order(-genecol_sum)[1]]
				warning(paste("More than one possible gene column found in gene2cat, using the one headed",colnames(gene2cat)[genecol]))
			}
			if(length(genecol)==0){
				genecol=1
				warning(paste("Gene column could not be identified in gene2cat conclusively, using the one headed",colnames(gene2cat)[genecol]))
			}
			othercol=1
			if(genecol==1){othercol=2}
			#Now put it into our delicious listy format
			gene2cat=split(gene2cat[,othercol],gene2cat[,genecol])
			#Do the appropriate builds
			cat2gene=reversemapping(gene2cat)
			gene2cat=reversemapping(cat2gene)
		}
		#!!!!
		#The following conditional has been flagged as a potential issue when using certain 
		#types of input where the category names are the same as gene names (which seems like 
		#something you should avoid anyway...).  Leave it for now
		#!!!!
		#We're now garunteed to have a list (unless the user screwed up the input) but it could 
		#be category->genes rather than the gene->categories that we want. 
		if(sum(unique(unlist(gene2cat,use.names=FALSE))%in%rownames(pwf))>sum(unique(names(gene2cat))%in%rownames(pwf))){
			gene2cat=reversemapping(gene2cat)
		}
		#Alright, we're garunteed a list going in the direction we want now.  Throw out genes which we will not use
		gene2cat=gene2cat[names(gene2cat)%in%rownames(pwf)]

		#Rebuild because it's a fun thing to do
		cat2gene=reversemapping(gene2cat)
		gene2cat=reversemapping(cat2gene)

		## make sure we remove duplicate entries .. e.g. see 
		## http://permalink.gmane.org/gmane.science.biology.informatics.conductor/46876
		cat2gene=lapply(cat2gene,function(x){unique(x)})
		gene2cat=lapply(gene2cat,function(x){unique(x)})
	}

	nafrac=(sum(is.na(pwf$pwf))/nrow(pwf))*100
	if(nafrac>50){
		warning(paste("Missing length data for ",round(nafrac),"% of genes.  Accuarcy of GO test will be reduced.",sep=''))
	}
	#Give the genes with unknown length the weight used by the median gene (not the median weighting!)
	pwf$pwf[is.na(pwf$pwf)]=pwf$pwf[match(sort(pwf$bias.data[!is.na(pwf$bias.data)])[ceiling(sum(!is.na(pwf$bias.data))/2)],pwf$bias.data)]

	###################### Calculating the p-values ########################
	# Remove all the genes with unknown GOterms
	unknown_go_terms=nrow(pwf)-length(gene2cat)
	if((!use_genes_without_cat) && unknown_go_terms>0 ){
	   message(paste("For",unknown_go_terms,"genes, we could not find any categories. These genes will be excluded."))
	   message("To force their use, please run with use_genes_without_cat=TRUE (see documentation).")
	   message("This was the default behavior for version 1.15.1 and earlier.")
	   pwf=pwf[rownames(pwf) %in% names(gene2cat),]
	} 
	#A few variables are always useful so calculate them
	cats=names(cat2gene)
	DE=rownames(pwf)[pwf$DEgenes==1]
	num_de=length(DE)
	num_genes=nrow(pwf)
	pvals=data.frame(category=cats,over_represented_pvalue=NA,under_represented_pvalue=NA,stringsAsFactors=FALSE,numDEInCat=NA,numInCat=NA)
	if(method=="Sampling"){
		#We need to know the number of DE genes in each category, make this as a mask that we can use later...
		num_DE_mask=rep(0,length(cats))
		a=table(unlist(gene2cat[DE],FALSE,FALSE))
		num_DE_mask[match(names(a),cats)]=as.numeric(a)
		num_DE_mask=as.integer(num_DE_mask)
		#We have to ensure that genes not associated with a category are included in the simulation, to do this they need an empty entry in the gene2cat list
		gene2cat=gene2cat[rownames(pwf)]
		names(gene2cat)=rownames(pwf)
		message("Running the simulation...")
		#Now do the actual simulating
		lookup=matrix(0,nrow=repcnt,ncol=length(cats))
		for(i in 1:repcnt){
			#A more efficient way of doing weighted random sampling without replacment than the built in function
			#The order(runif...)[1:n] bit picks n genes at random, weighting them by the PWF
			#The table(as.character(unlist(...))) bit then counts the number of times this random set occured in each category
			a=table(as.character(unlist(gene2cat[order(runif(num_genes)^(1/pwf$pwf),decreasing=TRUE)[1:num_de]],FALSE,FALSE)))
			lookup[i,match(names(a),cats)]=a
			pp(repcnt)
		}
		message("Calculating the p-values...")
		#The only advantage of the loop is it uses less memory...
		#for(i in 1:length(cats)){
		#	pvals[i,2:3]=c((sum(lookup[,i]>=num_DE_mask[i])+1)/(repcnt+1),(sum(lookup[,i]<=num_DE_mask[i])+1)/(repcnt+1))
		#	pp(length(cats))
		#}
		pvals[,2]=(colSums(lookup>=outer(rep(1,repcnt),num_DE_mask))+1)/(repcnt+1)
		pvals[,3]=(colSums(lookup<=outer(rep(1,repcnt),num_DE_mask))+1)/(repcnt+1)
	}
	if(method=="Wallenius"){
		message("Calculating the p-values...")
		#All these things are just to make stuff run faster, mostly because comparison of integers is faster than string comparison
		degenesnum=which(pwf$DEgenes==1)
		#Turn all genes into a reference to the pwf object
		cat2genenum=relist(match(unlist(cat2gene),rownames(pwf)),cat2gene)
		#This value is used in every calculation, by storing it we need only calculate it once
		alpha=sum(pwf$pwf)
		#Each category will have a different weighting so needs its own test
		pvals[,2:3]=t(sapply(cat2genenum,function(u){
			#The number of DE genes in this category
			num_de_incat=sum(degenesnum%in%u)
			#The total number of genes in this category
			num_incat=length(u)
			#This is just a quick way of calculating weight=avg(PWF within category)/avg(PWF outside of category)
			avg_weight=mean(pwf$pwf[u])
			weight=(avg_weight*(num_genes-num_incat))/(alpha-num_incat*avg_weight)
			if(num_incat==num_genes){ weight=1 } #case for the root GO terms			
			#Now calculate the sum of the tails of the Wallenius distribution (the p-values)
			c(dWNCHypergeo(num_de_incat,num_incat,num_genes-num_incat,num_de,weight)
			+pWNCHypergeo(num_de_incat,num_incat,num_genes-num_incat,num_de,weight,lower.tail=FALSE),
			pWNCHypergeo(num_de_incat,num_incat,num_genes-num_incat,num_de,weight))
		}))
	}
	if(method=="Fishers"){
		message("Calculating the p-values...")
		#All these things are just to make stuff run faster, mostly because comparison of integers is faster than string comparison
		degenesnum=which(pwf$DEgenes==1)
		#Turn all genes into a reference to the pwf object
		cat2genenum=relist(match(unlist(cat2gene),rownames(pwf)),cat2gene)
		#This value is used in every calculation, by storing it we need only calculate it once
		alpha=sum(pwf$pwf)
		#Each category will have a different weighting so needs its own test
		pvals[,2:3]=t(sapply(cat2genenum,function(u){
			#The number of DE genes in this category
			num_de_incat=sum(degenesnum%in%u)
			#The total number of genes in this category
			num_incat=length(u)
			#This is just a quick way of calculating weight=avg(PWF within category)/avg(PWF outside of category)
			avg_weight=mean(pwf$pwf[u])
			weight=(avg_weight*(num_genes-num_incat))/(alpha-num_incat*avg_weight)
			if(num_incat==num_genes){ weight=1 } #case for the root GO terms			
			#Now calculate the sum of the tails of the Wallenius distribution (the p-values)
			c(dFNCHypergeo(num_de_incat,num_incat,num_genes-num_incat,num_de,weight)
			+pFNCHypergeo(num_de_incat,num_incat,num_genes-num_incat,num_de,weight,lower.tail=FALSE),
			pFNCHypergeo(num_de_incat,num_incat,num_genes-num_incat,num_de,weight))
		}))
	}
	if(method=="Hypergeometric"){
		message("Calculating the p-values...")
		#All these things are just to make stuff run faster, mostly because comparison of integers is faster than string comparison
		degenesnum=which(pwf$DEgenes==1)
		#Turn all genes into a reference to the pwf object
		cat2genenum=relist(match(unlist(cat2gene),rownames(pwf)),cat2gene)
		#Simple hypergeometric test, one category at a time
		pvals[,2:3]=t(sapply(cat2genenum,function(u){
			#The number of DE genes in this category
			num_de_incat=sum(degenesnum%in%u)
			#The total number of genes in this category
			num_incat=length(u)
			#Calculate the sum of the tails of the hypergeometric distribution (the p-values)
			c(dhyper(num_de_incat,num_incat,num_genes-num_incat,num_de)+phyper(num_de_incat,num_incat,num_genes-num_incat,num_de,lower.tail=FALSE),phyper(num_de_incat,num_incat,num_genes-num_incat,num_de))
		}))
	}
   #Populate the count columns...
   degenesnum=which(pwf$DEgenes==1)
   cat2genenum=relist(match(unlist(cat2gene),rownames(pwf)),cat2gene)
   pvals[,4:5]=t(sapply(cat2genenum,function(u){
      c(sum(degenesnum%in%u),length(u))
   }))
   
   #Finally, sort by p-value
   pvals=pvals[order(pvals$over_represented_pvalue),]

   # Supplement the table with the GO term name and ontology group
   # but only if the enrichment categories are actually GO terms
   if(any(grep("^GO:",pvals$category))){
      GOnames=select(GO.db,keys=pvals$category,columns=c("TERM","ONTOLOGY"))[,2:3]
      colnames(GOnames)<-tolower(colnames(GOnames))
      pvals=cbind(pvals,GOnames)
   }

   # And return
   return(pvals)
}
```



```{r}
columns(org.Hs.eg.db)
```
```{r}
keys(org.Hs.eg.db)
```



