---
title: "04_compareGeneOntology"
author: "Caitlin Page"
date: "2024-09-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction
```{r}
library(ggVennDiagram)
```

```{r}
GO.wall <- readRDS("../output/goseq_res.rds")
GO.wall.null <- readRDS("../output/goseq_null_res.rds")

mm_hack <- readRDS("../output/mm_hack_res.rds")

sig_terms_array <- readRDS("../output/mm_array_res.rds")
sig_terms_array_biased <- readRDS("../output/mm_array_null.rds")
```

# Compare the significant terms between approaches

## Venn diagrams
```{r}
list_venn <- list(bias = GO.wall.null$term, correct = GO.wall$term)
ggVennDiagram(list_venn)

list_venn <- list(bias = GO.wall.null$term, correct = mm_hack$TERM)
ggVennDiagram(list_venn)
```
```{r}
list_venn <- list(correct = sig_terms_array$TERM, bias = sig_terms_array_biased$TERM)
ggVennDiagram(list_venn)
```
```{r}
list_venn <- list(seq_correct = GO.wall$term, array_correct = sig_terms_array$TERM)
ggVennDiagram(list_venn)
```
```{r}
list_venn <- list(seq_correct = GO.wall$term, array_bias = sig_terms_array_biased$TERM)
ggVennDiagram(list_venn)
```
```{r}
list_venn <- list(seq_bias = GO.wall.null$term, array_correct = sig_terms_array$TERM)
ggVennDiagram(list_venn)
```
```{r}
list_venn <- list(seq_bias = GO.wall.null$term, array_bias = sig_terms_array_biased$TERM)
ggVennDiagram(list_venn)
```


```{r}
list_venn <- list(array_correct = sig_terms_array$TERM, array_bias = sig_terms_array_biased$TERM, seq_correct = GO.wall$term, seq_bias = GO.wall.null$term)
ggVennDiagram(list_venn)
```
## Upset plot



####
# terms no longer signif - if they match up
```{r}
go_entrez %>% distinct(go, array_cat_signif, array_cat_signif_bias, seq_cat_signif, seq_cat_signif_bias) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_sig_a = ifelse(array_cat_signif == FALSE, TRUE, FALSE))
```
```{r}
go_entrez %>% distinct(go, array_cat_signif, array_cat_signif_bias, seq_cat_signif, seq_cat_signif_bias) %>% filter(seq_cat_signif_bias == TRUE) %>% mutate(stop_sig_s = ifelse(seq_cat_signif == FALSE, TRUE, FALSE)) 
```
```{r}
ggVennDiagram(list(array_not = go_entrez %>% distinct(go, array_cat_signif, array_cat_signif_bias, seq_cat_signif, seq_cat_signif_bias) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_sig_a = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>% filter(stop_sig_a == TRUE) %>% .$go, 
              seq_not = go_entrez %>% distinct(go, array_cat_signif, array_cat_signif_bias, seq_cat_signif, seq_cat_signif_bias) %>% filter(seq_cat_signif_bias == TRUE) %>% mutate(stop_sig_s = ifelse(seq_cat_signif == FALSE, TRUE, FALSE)) %>% filter(stop_sig_s == TRUE) %>% .$go))
```
- ok so that doesn't agree

-maybe categories that become significant?
-so filter to the false and then it's whatever the unbiased picks up as true
```{r}
ggVennDiagram(list(array_not = go_entrez %>% distinct(go, array_cat_signif, array_cat_signif_bias, seq_cat_signif, seq_cat_signif_bias) %>% filter(array_cat_signif_bias == FALSE) %>% mutate(stop_sig_a = ifelse(array_cat_signif == TRUE, TRUE, FALSE)) %>% filter(stop_sig_a == TRUE) %>% .$go, 
              seq_not = go_entrez %>% distinct(go, array_cat_signif, array_cat_signif_bias, seq_cat_signif, seq_cat_signif_bias) %>% filter(seq_cat_signif_bias == FALSE) %>% mutate(stop_sig_s = ifelse(seq_cat_signif == TRUE, TRUE, FALSE)) %>% filter(stop_sig_s == TRUE) %>% .$go))
```
-yeah that's a no as expected - array all the unbiased is within the biased approach
##############################

```{r}
mm_hack
```

## more complicated stuff
compare number of genes with many probes in sets without bias correction compared to with bias correction.
So the categories that are significant before correction might have higher than average CpG content...

but the result we have is terms and genes
and i can't get the genes back
```{r}
sig_terms_array
```
- so i want some way to add the genes back in?
- if i can match entrez ids to go terms i'm good - can go from there

```{r}
library(org.Hs.eg.db)
```

```{r}
x <- org.Hs.egGO
# Get the entrez gene identifiers that are mapped to a GO ID
entrez_map <- mappedkeys(x)
# Convert to a list
xx <- as.list(x[entrez_map])
```
```{r}
xx[[1]][[1]][["GOID"]]
str(xx)
```
```{r}
entrez_map[1:10]
length(entrez_map)
```
- that was the exact match version
- so like one gene per go
- we don't want that

```{r}
# For org.Hs.egGO2ALLEGS
xx <- as.list(org.Hs.egGO2ALLEGS)
if(length(xx) > 0){
    # Gets the Entrez Gene identifiers for the top 2nd and 3nd GO identifiers
    goids <- xx[2:3]
    # Gets all the Entrez Gene identifiers for the first element of goids
    goids[[1]]
    # Evidence code for the mappings
    names(goids[[1]])
}
```
i don't have the brainspace for this
ok so xx is a list 
each section is a go term, and in it, is the entrez ids of genes in it
```{r}
names(xx)[1:3]
```



and here I have the entrez gene ids and the number of cpg sites
i want a way to pass over the go term to this?
maybe it would be easier to pass it over to the results which has the go terms
```{r}
sig_terms_array %>% mutate(test_pass = ifelse(rownames(.) %in% names(xx), TRUE, FALSE))
```
- ok this is the "set"
- now just need the biases to go with
- i could just do a massive unlist - i think
-actually that worked - except for the names bit
```{r}
names(unlist(xx))[1:5]
```

```{r}
go_entrez <- data.frame(cbind(go = gsub("\\..*", "", names(unlist(xx))), entrez = unlist(xx)))
```
```{r}
go_entrez <- go_entrez %>% mutate(num_cg = data.frame(mapped_genes$freq)[match(.$entrez, data.frame(mapped_genes$freq)$eg.all), "Freq"])
go_entrez
```
```{r}
mapped_genes$freq %>% data.frame() %>% filter(eg.all == "28472")
sig_terms_array["GO:0002376",]
```
```{r}
unique(go_entrez$entrez) %in% mapped_genes$universe %>% summary()
mapped_genes$universe %in% unique(go_entrez$entrez) %>% summary()
```
- ok this also explains problems - they don't match
- is it that the missing ones have 0 cpgs??? - because those were removed in the mapped_genes thing it didn't want them

# let's limit to the matching ones for now
```{r}
go_entrez 
```

```{r}
go_entrez %>% filter(is.na(num_cg))
```

```{r}
go_entrez %>% distinct(go)
go_entrez %>% distinct(entrez)
go_entrez %>% distinct(num_cg)
```

```{r}
go_entrez <- go_entrez %>% mutate(array_cat_signif = ifelse(go %in% rownames(sig_terms_array), TRUE, FALSE),
                     array_cat_signif_bias = ifelse(go %in% rownames(sig_terms_array_biased), TRUE, FALSE),
                     seq_cat_signif = ifelse(go %in% GO.wall$category, TRUE, FALSE),
                     seq_cat_signif_bias = ifelse(go %in% GO.wall.null$category, TRUE, FALSE))
```
- ok I'm actually getting close now

```{r}
go_entrez %>% group_by(go) %>% mutate(num_entrez = dplyr::n(), median_cg = median(num_cg)) %>% distinct(.)
```
```{r}
go_entrez %>% group_by(go) %>% mutate(num_entrez = dplyr::n(), median_cg = median(num_cg)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, num_entrez, median_cg) %>% filter(array_cat_signif == TRUE) %>%
  ggplot(aes(x = median_cg, colour = array_cat_signif)) +
  geom_density()

go_entrez %>% group_by(go) %>% filter(array_cat_signif == TRUE) %>%
  ggplot(aes(x = num_cg, colour = array_cat_signif)) +
  geom_density()
```
-second plot i think is close?
except plot looks the same between signif and not signif
```{r}
go_entrez %>% group_by(go) %>% mutate(num_entrez = dplyr::n(), median_cg = median(num_cg)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, num_entrez, median_cg) %>% filter(array_cat_signif_bias == TRUE) %>%
  ggplot(aes(x = median_cg, colour = array_cat_signif_bias)) +
  geom_density()

go_entrez %>% group_by(go) %>% filter(array_cat_signif_bias == TRUE) %>%
  ggplot(aes(x = num_cg, colour = array_cat_signif_bias)) +
  geom_density()
```
hold up no we want to be comparing number of genes with many probes (many cpgs)
i'm so done
```{r}
go_entrez[,1:5] %>% melt(id = c("go", "entrez", "num_cg")) %>% filter(value == TRUE) %>%
  ggplot(aes(x = variable, y = num_cg)) +
  geom_boxplot()
```
- again they look the same
```{r}
go_entrez[,1:5] %>% melt(id = c("go", "entrez", "num_cg")) %>% filter(value == TRUE) %>% group_by(variable) %>% distinct(go) %>% summarise(dplyr::n())
```
```{r}
go_entrez[,1:5] %>% melt(id = c("go", "entrez", "num_cg")) %>% filter(value == TRUE) %>% distinct(.)
```

```{r}
go_entrez[,1:5] %>% melt(id = c("go", "entrez", "num_cg")) %>% filter(value == TRUE) %>% distinct(.) %>% .[1:500,] %>%
  ggplot(aes(x = go, colour = variable, y = num_cg)) +
  geom_boxplot()
```
```{r}
go_entrez[,1:5] %>% melt(id = c("go", "entrez", "num_cg")) %>% filter(value == TRUE) %>% distinct(.) %>% group_by(go, variable) %>% summarise(mean = mean(num_cg), median = median(num_cg)) 
```

```{r}
go_entrez[,1:5] %>% melt(id = c("go", "entrez", "num_cg")) %>% filter(value == TRUE) %>% distinct(.)
```
```{r}
go_entrez[,1:5] %>% filter(array_cat_signif == TRUE) %>% group_by(go) %>% summarise(mean = mean(num_cg), median = median(num_cg))
go_entrez[,1:5] %>% filter(array_cat_signif_bias == TRUE) %>% group_by(go) %>% summarise(mean = mean(num_cg), median = median(num_cg))
```
- ok so these numbers are diff now but also they are very similar
```{r}
go_entrez %>% filter(array_cat_signif == TRUE) %>%
  ggplot(aes(x = num_cg)) +
  geom_density()

go_entrez %>% filter(array_cat_signif_bias == TRUE) %>%
  ggplot(aes(x = num_cg)) +
  geom_density()

go_entrez %>% filter(array_cat_signif == TRUE) %>%
  ggplot(aes(x = log2(num_cg))) +
  geom_density()

go_entrez %>% filter(array_cat_signif_bias == TRUE) %>%
  ggplot(aes(x = log2(num_cg))) +
  geom_density()
```
```{r}
go_entrez %>% group_by(num_cg) %>% filter(is.na(num_cg))
```

```{r}
?mean
go_entrez[,1:5] %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE), median=median(num_cg, na.rm=TRUE), sd = sd(num_cg, na.rm=TRUE))
summary(go_entrez$num_cg)
go_entrez %>% 
  ggplot(aes(x = num_cg)) +
  geom_density()
```
mean of it all: 27
median: 36.47
- feels weird that max is 51 - feel like it should be higher

- let's get mean and median for each - and compare it to the overall
so like a z test basically

- let's try z test
- i think non-param version is chi-square

z need mu and sd of pop
27 & 38.39442
want for each of "sets" - signif go terms
```{r}
library(BSDA)
```

```{r}
list_go <- go_entrez[,1:5] %>% filter(array_cat_signif == TRUE) %>% distinct(go) %>% .$go
list_go[1:5]
length(list_go)

res_stat <- c()
res_p <- c()
for(i in 1:73) {
  test <- z.test(filter(go_entrez, array_cat_signif == TRUE, go == list_go[i])$num_cg, mu = 27, sigma.x=38.39442)
  res_stat <- c(res_stat, test$statistic)
  res_p <- c(res_p, test$p.value)
}
array_signif_z <- cbind(go=list_go, z_stat = res_stat, p_val = res_p) %>% data.frame()
array_signif_z 
```
```{r}
array_signif_z %>% mutate(z_stat = as.double(z_stat), p_val = as.double(p_val)) %>% filter(p_val < 0.05)
array_signif_bias_z %>% mutate(z_stat = as.double(z_stat), p_val = as.double(p_val)) %>% filter(p_val < 0.05)
```
- nothing signif
```{r}
list_go <- go_entrez[,1:5] %>% filter(array_cat_signif_bias == TRUE) %>% distinct(go) %>% .$go
list_go[1:5]

res_stat <- c()
res_p <- c()
for(i in 1:length(list_go)) {
  test <- z.test(filter(go_entrez, array_cat_signif_bias == TRUE, go == list_go[i])$num_cg, mu=27, sigma.x=38.39442)
  res_stat <- c(res_stat, test$statistic)
  res_p <- c(res_p, test$p.value)
}
array_signif_bias_z <- cbind(go=list_go, z_stat = res_stat, p_val = res_p) %>% data.frame()
array_signif_bias_z
```
```{r}
go_entrez %>% filter(array_cat_signif_bias == TRUE) %>% summarise(mean = mean(num_cg), sd(num_cg))
```
- doesn't change the stats


# what about a t test?

```{r}
t.test(filter(go_entrez, array_cat_signif == TRUE, array_cat_signif_bias == TRUE)$num_cg, 
       filter(go_entrez, array_cat_signif == FALSE, array_cat_signif_bias == TRUE)$num_cg)
```
- nope

# what about just the number of genes with more than average num_cg??
av = 27
```{r}
array_sig_num <- go_entrez[,1:5] %>% filter(array_cat_signif == TRUE) %>% group_by(go) %>% mutate(num_genes = dplyr::n()) %>% filter(num_cg > 27) %>% mutate(num_genes_above = dplyr::n(), ratio = dplyr::n()/num_genes) %>% distinct(go, num_genes, num_genes_above, ratio) %>% ungroup() %>% data.frame()
```
```{r}
array_bias_num <- go_entrez[,1:5] %>% filter(array_cat_signif_bias == TRUE) %>% group_by(go) %>% mutate(num_genes = dplyr::n()) %>% filter(num_cg > 27) %>% mutate(num_genes_above = dplyr::n(), ratio = dplyr::n()/num_genes) %>% distinct(go, num_genes, num_genes_above, ratio) %>% ungroup() %>% data.frame()
```

```{r}
array_bias_num %>% mutate(sig_ratio = array_sig_num[match(.$go, array_sig_num$go), "ratio"]) %>% filter(!is.na(sig_ratio)) %>%
  ggplot(aes(x = ratio, y = sig_ratio)) +
  geom_point()
```

# alicia says average should be average per category

```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg)

go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% 
  ggplot(aes(y = mean_cg)) +
  geom_boxplot()

go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif == TRUE) %>% 
  ggplot(aes(y = mean_cg)) +
  geom_boxplot()
```

```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif == TRUE | array_cat_signif_bias == TRUE) %>% melt(id = c("go", "mean_cg")) %>% filter(value == TRUE) %>%
  ggplot(aes(x = variable, y = mean_cg)) +
  geom_boxplot()
```
```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(array_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category")
```
- ok now I think I'm rightish
- ok now we're talking

alicia wants to know how many categories in each box
```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>% ungroup() %>% group_by(stop_signif) %>% summarise(dplyr::n())
```
```{r}
go_entrez[,1:5] %>% filter(array_cat_signif == TRUE | array_cat_signif_bias == TRUE) %>% mutate(median = median(num_cg, na.rm=TRUE))
```
```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>% filter(stop_signif == TRUE) %>% .$mean_cg %>% summary()

go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>% filter(stop_signif == FALSE) %>% .$mean_cg %>% summary()
```
```{r}
t.test(go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>% filter(stop_signif == TRUE) %>% .$mean_cg, go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>% filter(stop_signif == FALSE) %>% .$mean_cg)
```



- try it as a density
```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, mean_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>%
  ggplot(aes(colour = stop_signif, x = mean_cg)) +
  geom_density()
```
- yeah they are the same
-try median

```{r}
go_entrez[,1:5] %>% group_by(go) %>% mutate(median_cg = median(num_cg, na.rm=TRUE)) %>% distinct(go, array_cat_signif, array_cat_signif_bias, median_cg) %>% filter(array_cat_signif_bias == TRUE) %>% mutate(stop_signif = ifelse(array_cat_signif == FALSE, TRUE, FALSE)) %>%
  ggplot(aes(x = stop_signif, y = median_cg)) +
  geom_boxplot()
```

########
## ok this boxplot thing but for sequencing

```{r}
go_entrez %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)"))
```


```{r}
go_entrez %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category")

go_entrez %>% distinct() %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category")
```
- alicia was hoping it would be clearer

-maybe median?

```{r}
go_entrez %>% group_by(go) %>% mutate(median_cg = median(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, median_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = median_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category")
```
- no not really

-maybe this is the NA problem?

```{r}
go_entrez %>% distinct() %>% mutate(num_cg_na = ifelse(is.na(num_cg), TRUE, FALSE)) %>% group_by(go) %>% mutate(num_gene = dplyr::n()) %>% ungroup() %>%
  group_by(go, num_gene, num_cg_na) %>% summarise(n=dplyr::n(), ratio = dplyr::n()/num_gene) %>% distinct() 
go_entrez %>% distinct() %>% filter(seq_cat_signif == TRUE | seq_cat_signif_bias == TRUE) %>% mutate(num_cg_na = ifelse(is.na(num_cg), TRUE, FALSE)) %>% group_by(go) %>% mutate(num_gene = dplyr::n()) %>% ungroup() %>%
  group_by(go, num_gene, num_cg_na) %>% summarise(n=dplyr::n(), ratio = dplyr::n()/num_gene) %>% distinct() %>% filter(num_cg_na == TRUE) %>% .[order(.$ratio, decreasing = TRUE),]
```
```{r}
GO.wall %>% filter(category == "GO:0019814")
```
# should be using the goseq matching probably - maybe that will make it look more elegant
```{r}
gene2cat <- goseq::getgo(test_bias$ensembl_gene_id, "hg19", "ensGene")
names(gene2cat) <- test_bias$ensembl_gene_id
#Do the two rebuilds to remove any nulls
reversemapping=function(map){
	tmp=unlist(map,use.names=FALSE)
	names(tmp)=rep(names(map),times=as.numeric(summary(map)[,1]))
	return(split(names(tmp),as.vector(tmp)))
}
cat2gene <- reversemapping(gene2cat)
gene2cat <- reversemapping(cat2gene)
```
```{r}
if(sum(unique(unlist(gene2cat,use.names=FALSE))%in%test_bias$ensembl_gene_id)>sum(unique(names(gene2cat))%in%test_bias$ensembl_gene_id)){
			gene2cat <- reversemapping(gene2cat)
		}
		#Alright, we're garunteed a list going in the direction we want now.  Throw out genes which we will not use
		gene2cat <- gene2cat[names(gene2cat)%in%test_bias$ensembl_gene_id]

		#Rebuild because it's a fun thing to do
		cat2gene <- reversemapping(gene2cat)
		gene2cat <- reversemapping(cat2gene)

		## make sure we remove duplicate entries .. e.g. see 
		## http://permalink.gmane.org/gmane.science.biology.informatics.conductor/46876
		cat2gene <- lapply(cat2gene,function(x){unique(x)})
		gene2cat <- lapply(gene2cat,function(x){unique(x)})
```


```{r}
head(cat2gene)[1]
```
```{r}
GO.wall
```

```{r}
cat2gene2 <- cat2gene
names(cat2gene2) <- paste0(names(cat2gene2), ".")
go_ensembl <- data.frame(cbind(go = gsub("\\..*", "", names(unlist(cat2gene2))), ensembl = unlist(cat2gene)))
go_ensembl %>% distinct(ensembl)
go_ensembl %>% distinct(go)
```
```{r}
go_ensembl %>% distinct()
```



```{r}
unique(go_entrez$go) %in% unique(go_ensembl$go) %>% summary()
```

```{r}
unique(go_ensembl$ensembl) %in% test_bias$ensembl_gene_id %>% summary()
GO.wall$category %in% unique(go_ensembl$go) %>% summary()
test_bias <- data.frame(test_bias)
test_bias
GO.wall
GO.wall.null
go_ensembl
```



```{r}
try_res <- data.frame(category = names(cat2gene), numDEInCat = NA, numInCat=NA)
degenesnum=which(test_bias$has_dmr==1)
try_res[,2:3]=t(sapply(cat2genenum,function(u){
      c(sum(degenesnum%in%u),length(u))
   }))
```
```{r}
go_ensembl %>% filter(go == "GO:0000002")
"GO:0000002" %in% names(cat2gene)
```

```{r}
go_ensembl <- go_ensembl %>% mutate(num_cg = test_bias[match(.$ensembl, test_bias$ensembl_gene_id), "num_cg"],
                      gene_signif = test_bias[match(.$ensembl, test_bias$ensembl_gene_id), "has_dmr"],
                      gene_signif = ifelse(gene_signif == 0, FALSE, TRUE),
                      seq_cat_signif = ifelse(go %in% GO.wall$category, TRUE, FALSE),
                      seq_cat_signif_bias = ifelse(go %in% GO.wall.null$category, TRUE, FALSE))
```

```{r}
go_ensembl %>% distinct() %>% group_by(go) %>% mutate(median_cg = median(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, median_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = median_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Median number of CpGs per ontology category")

go_ensembl %>% distinct() %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category")
```
- still doesn't look as good as we hoped it would
```{r}
go_ensembl %>% distinct() %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category") +
  theme(legend.position = "none")
```

```{r}
png(filename="../output/boxplot.png", res=5000) # perhaps width/height as well
plot <- go_ensembl %>% distinct() %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category") +
  theme(legend.position = "none")
plot
dev.off()
```

- oops i think I've done it wrong
wait no
this plot is more about what we lose rather than gaining stuff
```{r}
go_ensembl %>% distinct() %>% group_by(go) %>% mutate(mean_cg = mean(num_cg, na.rm=TRUE)) %>% distinct(go, seq_cat_signif, seq_cat_signif_bias, mean_cg) %>% filter(seq_cat_signif_bias == TRUE) %>% 
  mutate(Method = ifelse(seq_cat_signif == FALSE, "Biased (Hypergeometric)", "Bias corrected (Wallenius)")) %>%
  ggplot(aes(x = Method, y = mean_cg, fill = Method)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Average number of CpGs per ontology category")
```


```{r}
go_entrez %>% distinct() %>% mutate(num_cg_na = ifelse(is.na(num_cg), TRUE, FALSE)) %>% group_by(go) %>% mutate(num_gene = dplyr::n()) %>% ungroup() %>%
  group_by(go, num_gene, num_cg_na) %>% summarise(n=dplyr::n(), ratio = dplyr::n()/num_gene) %>% distinct() %>% filter(num_cg_na == TRUE) %>%
  ggplot(aes(x = num_gene, y = n)) +
  geom_point()
```

```{r}
GO.wall %>% filter(category %in% filter(go_entrez, is.na(num_cg), seq_cat_signif == TRUE)$go)
GO.wall %>% filter(category %in% filter(go_entrez, is.na(num_cg), seq_cat_signif_bias == TRUE)$go)
```
- let's not use the entrez freq maybe
- let's do the 



- an upset plot would also be nice
