---
title: "04_compareSteps"
author: "Caitlin Page"
date: "2025-02-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

```{r}
library(plyranges)
library(dplyr)
library(ggplot2)
library(ggVennDiagram)
```


## DM Analysis results
```{r}
anno_array <- readRDS("../output/anno_array.rds")
array_dmps <- readRDS("../output/array_dmps.rds")
```

```{r}
bsseq_res <- readRDS("../output/bsseq_res.rds")
bsseq_dmrs <- readRDS("../output/bsseq_dmrs.rds")
```
```{r}
dmrcate_array_dmr <- dmrcate_array_dmr %>% data.frame()
dmrcate_array_dmr$rank <- 1:nrow(dmrcate_array_dmr)
```
```{r}
summary(dmrcate_array_dmr$min_smoothed_fdr)
summary(dmrcate_array_dmr$no.cpgs)
dmrcate_array_dmr %>% .[order(.$min_smoothed_fdr),]
```


## Compare indiv CpGs
* filter results 
** Bsseq doesn't have a PValue - individual CpGs are combined into DMRs based on their tstat 
** quantile cutoff (0.025, 0.975)
** (assuming it would do this on the t.stat.corrected values)
```{r}
bsseq_res
```
```{r}
quantile(bsseq_res$tstat.corrected, probs = c(0.025, 0.975))
```

```{r}
bsseq_res_filt <- bsseq_res %>% filter(tstat.corrected <= quantile(bsseq_res$tstat.corrected, probs = c(0.025, 0.975))[[1]] | 
                       tstat.corrected >= quantile(bsseq_res$tstat.corrected, probs = c(0.025, 0.975))[[2]])
bsseq_res_filt <- bsseq_res_filt %>% .[order(abs(.$tstat.corrected), decreasing = TRUE),] %>% mutate(rank = 1:n())
```
```{r}
nrow(bsseq_res)
nrow(array_dmps)
```
```{r}
compare_pos_res <- pair_nearest(as_granges(bsseq_res), as_granges(array_dmps)) %>% data.frame() %>% 
  mutate(signif_seq = ifelse(tstat.corrected <= quantile(bsseq_res$tstat.corrected, probs = c(0.025, 0.975))[[1]] | 
                       tstat.corrected >= quantile(bsseq_res$tstat.corrected, probs = c(0.025, 0.975))[[2]], TRUE, FALSE)) %>%
  mutate(signif_either = case_when(signif == TRUE & signif_seq == TRUE ~ "Both",
                                   signif == TRUE & signif_seq == FALSE ~ "Array",
                                   signif == FALSE & signif_seq == TRUE ~ "Seq",
                                   TRUE ~ "None"),
         gap_seq_minus_array = granges.x.start - granges.y.start)
```


```{r}
compare_pos_res %>%
  ggplot(aes(x = signif_either, y = gap_seq_minus_array, colour = signif_either)) +
  geom_boxplot()
```


```{r}
compare_pos_res %>%
  ggplot(aes(x = gap_seq_minus_array, colour = signif_either)) +
  geom_density()
```

```{r}
pair_nearest(as_granges(bsseq_res_filt), as_granges(array_dmps_filt)) %>% data.frame() %>% 
  mutate(gap_seq_minus_array = granges.x.start - granges.y.start) %>%
  ggplot(aes(x = gap_seq_minus_array)) +
  geom_density()

pair_nearest(as_granges(bsseq_res), as_granges(array_dmps)) %>% data.frame() %>% 
  mutate(gap_seq_minus_array = granges.x.start - granges.y.start) %>%
  ggplot(aes(x = gap_seq_minus_array)) +
  geom_density()
```

```{r}
actual_overlap_pos <- find_overlaps(as_granges(bsseq_res), as_granges(array_dmps), maxgap = 2) %>% data.frame()
ggVennDiagram(list(array_rank = array_dmps$rank, overlap = unique(actual_overlap_pos$rank)))
ggVennDiagram(list(seq_rank = bsseq_res$position, overlap = unique(actual_overlap_pos$position)))
```
* hold up - try overlap of signif just starting with the overlapping positions:
```{r}
ggVennDiagram(list(array_rank = array_dmps_filt$rank, overlap = unique(actual_overlap_pos$rank)))
ggVennDiagram(list(seq_rank = bsseq_res_filt$position, overlap = unique(actual_overlap_pos$position)))
```


```{r}
overlap_pos <- find_overlaps(as_granges(bsseq_res_filt), as_granges(array_dmps_filt), maxgap = 10) %>% data.frame()
overlap_pos
ggVennDiagram(list(array_rank = array_dmps_filt$array_rank, overlap = unique(overlap_pos$array_rank)))
ggVennDiagram(list(seq_rank = bsseq_res_filt$rank, overlap = unique(overlap_pos$rank.x)))

overlap_pos %>% 
  ggplot(aes(x = array_rank, y = rank.x)) +
  geom_point(alpha = 0.2)
```
```{r}
nrow(bsseq_res_filt)
nrow(array_dmps_filt)
nrow(overlap_pos)
```
```{r}
length(unique(overlap_pos$rank.x))
length(unique(overlap_pos$array_rank))
```

```{r}
overlap_pos

```

```{r}
overlap_pos
union_pos <- (as_granges(bsseq_res_filt), as_granges(array_dmps_filt), maxgap = 100) %>% data.frame()

ggVennDiagram(list(array_rank = array_dmps_filt$array_rank, overlap = unique(overlap_pos$array_rank)))
ggVennDiagram(list(seq_rank = bsseq_res_filt$rank, overlap = unique(overlap_pos$rank.x)))
```



## Overlap array sites and Bsseq DMR
```{r}
array_dmps_filt <- array_dmps %>% filter(adj.P.Val < 0.01)
bsseq_dmrs_filt <- bsseq_dmrs %>% filter(abs(areaStat) > 100)
```

```{r}
array_dmps_filt$array_rank <- 1:nrow(array_dmps_filt)
bsseq_dmrs_filt$seq_rank <- 1:nrow(bsseq_dmrs_filt)
```

```{r}
overlap <- plyranges::find_overlaps(plyranges::as_granges(bsseq_dmrs_filt), plyranges::as_granges(array_dmps_filt), maxgap = 100) %>% data.frame()

ggVennDiagram(list(array_rank = array_dmps_filt$array_rank, overlap = unique(overlap$array_rank)))
ggVennDiagram(list(seq_rank = bsseq_dmrs_filt$seq_rank, overlap = unique(overlap$seq_rank)))
```
* only 13% of the DMPs overlap with DMRs
* 56% of the DMRs overlap with DMPs
```{r}
overlap_10k <- plyranges::find_overlaps(plyranges::as_granges(bsseq_dmrs_filt), plyranges::as_granges(array_dmps_filt[1:10000,]), maxgap = 100) %>% data.frame()

ggVennDiagram(list(array_rank = array_dmps_filt$array_rank, overlap_10k = unique(overlap_10k$array_rank)))
ggVennDiagram(list(seq_rank = bsseq_dmrs_filt$seq_rank, overlap_10k = unique(overlap_10k$seq_rank)))
```
* 3% overlap
* 18% overlap
* adding maxgap=50 or 100 doesn't change stats

## Overlap DMRs
```{r}
bsseq_dmrs$bsseq_rank <- 1:nrow(bsseq_dmrs)
```

```{r}
overlap <- plyranges::find_overlaps(plyranges::as_granges(bsseq_dmrs), plyranges::as_granges(dmrcate_array_dmr), maxgap = 100) %>% data.frame()

overlap %>% 
  ggplot(aes(x = bsseq_rank, y = rank)) +
  geom_point(alpha = 0.3)

ggVennDiagram(list(array_rank = dmrcate_array_dmr$rank, overlap = unique(overlap$rank)))
ggVennDiagram(list(seq_rank = bsseq_dmrs$bsseq_rank, overlap = unique(overlap$bsseq_rank)))
```

```{r}
overlap <- plyranges::find_overlaps(plyranges::as_granges(bsseq_dmrs_filt), plyranges::as_granges(dmrcate_array_dmr), maxgap = 100) %>% data.frame()

overlap %>% 
  ggplot(aes(x = seq_rank, y = rank)) +
  geom_point(alpha = 0.3)

ggVennDiagram(list(array_rank = dmrcate_array_dmr$rank, overlap = unique(overlap$rank)))
ggVennDiagram(list(seq_rank = bsseq_dmrs_filt$seq_rank, overlap = unique(overlap$seq_rank)))
```


```{r}
overlap <- plyranges::find_overlaps(plyranges::as_granges(array_dmps), plyranges::as_granges(dmrcate_array_dmr), maxgap = 100) %>% data.frame()

ggVennDiagram(list(array_dmr = dmrcate_array_dmr$rank, overlap = unique(overlap$rank.y)))
ggVennDiagram(list(array_rank = array_dmps$rank, overlap = unique(overlap$rank.x)))
```

```{r}
overlap <- plyranges::find_overlaps(plyranges::as_granges(array_dmps_filt), plyranges::as_granges(dmrcate_array_dmr), maxgap = 100) %>% data.frame()

ggVennDiagram(list(array_dmr = dmrcate_array_dmr$rank, overlap = unique(overlap$rank.y)))
ggVennDiagram(list(array_rank = array_dmps_filt$rank, overlap = unique(overlap$rank.x)))
```

```{r}
overlap <- plyranges::find_overlaps(plyranges::as_granges(array_dmps_filt[1:10000,]), plyranges::as_granges(dmrcate_array_dmr), maxgap = 100) %>% data.frame()

ggVennDiagram(list(array_dmr = dmrcate_array_dmr$rank, overlap = unique(overlap$rank.y)))
ggVennDiagram(list(array_rank = array_dmps_filt[1:10000,]$rank, overlap = unique(overlap$rank.x)))
```


```{r}
array_dmps_filt
```
```{r}
overlap %>%
  ggplot(aes(x = array_rank, y = seq_rank)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 10000, colour = "red")

overlap_10k %>%
  ggplot(aes(x = array_rank, y = seq_rank)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 10000, colour = "red")
```
```{r}
overlap_10k %>%
  group_by(seq_rank) %>% filter(seq_rank <= 40) %>%
  ggplot(aes(x = as.factor(seq_rank), y = array_rank)) +
  geom_boxplot()
```
```{r}
overlap_10k %>%
  group_by(array_rank) %>% filter(array_rank <= 50) %>%
  ggplot(aes(x = as.factor(array_rank), y = seq_rank)) +
  geom_point()
```

* very little overlap
* does this follow through to the genes?

```{r}
overlap
```
```{r}
cor.test(overlap$array_rank, overlap$seq_rank)
cor.test(overlap_10k$array_rank, overlap_10k$seq_rank)
```


## compare genes

```{r}
array_genes <- getMappedEntrezIDs(rownames(array_dmps[1:10000,]), rownames(array_dmps))
```
```{r}
test_out_filt
```
```{r}
ggVennDiagram(list(uni_array = array_genes$universe, uni_seq = test_out_filt$universe))
ggVennDiagram(list(sig_array = array_genes$sig.eg, sig_seq = test_out_filt$sig.eg))
```
* little overlap
```{r}
cbind(entrez = mapped_entrez$sig.eg) %>% data.frame()
cbind(entrez = test_out_filt$sig.eg) %>% data.frame()
```

```{r}
genes_symbol %>% mutate(sig_array = ifelse(.$ENTREZID %in% mapped_entrez$sig.eg, TRUE, FALSE)) %>% group_by(sig_array) %>% summarise(dplyr::n())
genes_symbol %>% mutate(sig_seq = ifelse(.$ENTREZID %in% test_out_filt$sig.eg, TRUE, FALSE)) %>% group_by(sig_seq) %>% summarise(dplyr::n())
```
* rows match
```{r}
genes_symbol %>% mutate(sig_array = ifelse(.$ENTREZID %in% mapped_entrez$sig.eg, TRUE, FALSE),
                        sig_seq = ifelse(.$ENTREZID %in% test_out_filt$sig.eg, TRUE, FALSE)) %>% 
  mutate(signif = case_when(sig_array == FALSE & sig_seq == FALSE ~ "None",
                            sig_array == FALSE & sig_seq == TRUE ~ "Seq",
                            sig_array == TRUE & sig_seq == FALSE ~ "Array",
                            TRUE ~ "Both")) %>% 
  group_by(signif) %>% summarise(n = dplyr::n()) %>% ungroup() %>% mutate(prop = n/sum(n))
```

```{r}
genes_symbol %>% mutate(sig_array = ifelse(.$ENTREZID %in% mapped_entrez$sig.eg, TRUE, FALSE),
                        sig_seq = ifelse(.$ENTREZID %in% test_out_filt$sig.eg, TRUE, FALSE)) %>% 
  mutate(signif = case_when(sig_array == FALSE & sig_seq == FALSE ~ "None",
                            sig_array == FALSE & sig_seq == TRUE ~ "Seq",
                            sig_array == TRUE & sig_seq == FALSE ~ "Array",
                            TRUE ~ "Both")) %>% 
  filter(signif != "None") %>%
  group_by(signif) %>% summarise(n = dplyr::n()) %>% ungroup() %>% mutate(prop = n/sum(n))
```
```{r}
genes_both <- genes_symbol %>% mutate(sig_array = ifelse(.$ENTREZID %in% mapped_entrez$sig.eg, TRUE, FALSE),
                        sig_seq = ifelse(.$ENTREZID %in% test_out_filt$sig.eg, TRUE, FALSE)) %>%
  filter(sig_array == TRUE & sig_seq == TRUE)
```

## thoughts
* I think the samples were more different than we thought
* or I only analyse the overlap

```{r}
test_out_both <- run_miss_methyl_a(biomart_genes, filter(test_anno_filt, gene_name %in% genes_both$SYMBOL))
```

```{r}
test_res_both <- run_miss_methyl_b(test_out_both, "Wallenius")
```
* ok there is some immune stuff there in the 20s and 30s

```{r}
go_
```



