---
title: "05_notes"
author: "Caitlin Page"
date: "2025-01-31"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

```{r}
renv::dependencies() %>% data.frame()
```


```{r}
biomart_genes %>% mutate(width = end - start + 1, has_dmr = ifelse(ensembl_gene_id %in% filter(test_anno, abs(min_dist) == 0)$ensembl_gene_id, TRUE, FALSE)) %>%
  group_by(num_cg) %>%
    mutate(num_bins_same_cg = n()) %>% 
  .[order(.$num_cg),] %>%
  ungroup() %>%
  mutate(bin_group = rep(1:ceiling(nrow(biomart_genes)/100), each = 100)[1:nrow(biomart_genes)]) %>% 
  group_by(bin_group) %>%
  mutate(num_cg_bin_group = mean(num_cg)) %>%
  group_by(num_cg_bin_group) %>%
    mutate(num_bins_group_same_cg = dplyr::n()) %>%
    group_by(num_cg_bin_group, has_dmr) %>%
    mutate(num_per_dmr = dplyr::n(), prop = dplyr::n()/num_bins_group_same_cg) %>%
    distinct(num_cg_bin_group, num_bins_group_same_cg, num_per_dmr, prop, has_dmr) %>%
    group_by(num_cg_bin_group) %>%
    .[order(.$num_cg_bin_group),] %>%
    mutate(prop = ifelse(has_dmr == FALSE, 0, prop)) %>%
    mutate(num = dplyr::n()) %>%
    mutate(has_dmr = ifelse(num == 1, TRUE, has_dmr)) %>% filter(has_dmr == TRUE)
```

```{r}
bsseq_dmrs
```


```{r}
library(dplyr)
library(rbenchmark)
library(ggplot2)
library(ggVennDiagram)
```


STEPS:
```{r}
biomart_genes <- readRDS("../output/genes_biomart.rds")
bsseq_dmrs <- readRDS("../output/bsseq_dmrs.rds")
bsseq_res <- readRDS("../output/bsseq_res.rds")
```
```{r}
test_anno <- annoGeneDmr(bsseq_res, bsseq_dmrs, gene_source = "biomaRt", biomart_genes)
test_anno
```
```{r}
str(test_out_filt)
```



overall
```{r}
run_gst_seq(dmrs, tested_cpgs, 
                        gene_source = c("biomaRt", "ExperimentHub"), gene_feature=c("gene", "promoter"),
                        correct_bias=TRUE, plot_bias=TRUE, ...)
```
breakdown
```{r}
run_miss_methyl_a(biomart_genes, test_anno)
str(test_out_filt)
```
* coverts ensembl to entrez gene ids
* sets up for testing - de info
```{r}
go <- .getGO
```
```{r}
goseq_go <- getgo(biomart_genes$ensembl_gene_id, "hg19", "ensGene", "GO:BP")
```
```{r}
MM_array_wall <- readRDS("../output/mm_array_res.rds")
```
```{r}
MM_wall_filt
```


```{r}
ggVennDiagram(list(array_mm = rownames(MM_array_wall), seq_mm = filter(MM_wall_filt, FDR < 0.05)$GOID))
```

```{r}
go <- 
```

```{r}
benchmark(mm_a = run_miss_methyl_a(biomart_genes, test_anno_filt), replications = 50)
```

```{r}
run_miss_methyl_b(out, method = c("Wallenius", "Fishers", "None"), plot_bias=FALSE, ...)
```
* does the testing
* nature of the biasedUrn method is that it can't be // - each test must be done individually - so this is still slower then would be good
```{r}
benchmark(wall = run_miss_methyl_b(test_out_filt, method = "Wallenius"), fish = run_miss_methyl_b(test_out_filt, method = "Fishers"), replications = 5)
```
```{r}
# goseq version
benchmark(wall = goseq(pwf_seq, "hg19", "ensGene"), fish = goseq_mod(pwf_seq,"hg19","ensGene",method="Fishers"), replications = 50)
```
```{r}
GO.wall %>% filter(FDR >= 0.05)
GO.samp %>% filter(FDR >= 0.05)
GO.fish %>% filter(FDR >= 0.05)
```
* compare microarray ranking goseq
```{r}
GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = samp)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 1, colour = "blue")

GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 461, colour = "green")

GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = samp, y = fish)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 1, colour = "blue") +
  geom_hline(yintercept = 461, colour = "green")
```
```{r}
GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = samp)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 1, colour = "blue") +
  lims(x=c(0,1000),y=c(0,1000))

GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 461, colour = "green") +
  lims(x=c(0,1000),y=c(0,1000))

GO.wall %>% mutate(samp = GO.samp[match(GO.wall$category, GO.samp$category), "samp_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = samp, y = fish)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 1, colour = "blue") +
  geom_hline(yintercept = 461, colour = "green") +
  lims(x=c(0,1000),y=c(0,1000))
```

```{r}
GO.wall_seq %>% filter(FDR >= 0.05)
GO.fish_seq %>% filter(FDR >= 0.05)
```
* compare stat for WGBS goseq

```{r}
GO.wall_seq %>% mutate(fish_seq = GO.fish_seq[match(GO.wall_seq$category, GO.fish_seq$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 221, colour = "blue") +
  geom_hline(yintercept = 256, colour = "purple")

GO.wall_seq %>% mutate(fish_seq = GO.fish_seq[match(GO.wall_seq$category, GO.fish_seq$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 221, colour = "blue") +
  geom_hline(yintercept = 256, colour = "purple") +
  lims(x=c(0,1000),y=c(0,1000))
```
* compare microarray and WGBS ranking for goseq
```{r}
GO.wall %>% mutate(wall_seq = GO.wall_seq[match(GO.wall$category, GO.wall_seq$category), "wall_rank"],) %>%
  ggplot(aes(x = wall_rank, y = wall_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 221, colour = "blue") +
  lims(x=c(0,1000),y=c(0,1000))

GO.wall %>% mutate(fish_seq = GO.fish_seq[match(GO.wall$category, GO.fish_seq$category), "fish_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = fish, y = fish_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 461, colour = "green") +
  geom_hline(yintercept = 256, colour = "purple") +
  lims(x=c(0,1000),y=c(0,1000))

GO.wall %>% mutate(fish_seq = GO.fish_seq[match(GO.wall$category, GO.fish_seq$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 256, colour = "purple") +
  lims(x=c(0,1000),y=c(0,1000))

GO.wall %>% mutate(wall_seq = GO.wall_seq[match(GO.wall$category, GO.wall_seq$category), "wall_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = fish, y = wall_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 461, colour = "green") +
  geom_hline(yintercept = 221, colour = "blue") +
  lims(x=c(0,1000),y=c(0,1000))
```

```{r}
GO.wall %>% mutate(wall_seq = GO.wall_seq[match(GO.wall$category, GO.wall_seq$category), "wall_rank"],) %>%
  ggplot(aes(x = wall_rank, y = wall_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 221, colour = "blue") 


GO.wall %>% mutate(fish_seq = GO.fish_seq[match(GO.wall$category, GO.fish_seq$category), "fish_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = fish, y = fish_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 461, colour = "green") +
  geom_hline(yintercept = 256, colour = "purple") 


GO.wall %>% mutate(fish_seq = GO.fish_seq[match(GO.wall$category, GO.fish_seq$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 453, colour = "red") +
  geom_hline(yintercept = 256, colour = "purple") 


GO.wall %>% mutate(wall_seq = GO.wall_seq[match(GO.wall$category, GO.wall_seq$category), "wall_rank"],
                   fish = GO.fish[match(GO.wall$category, GO.fish$category), "fish_rank"]) %>%
  ggplot(aes(x = fish, y = wall_seq)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 461, colour = "green") +
  geom_hline(yintercept = 221, colour = "blue") 

```

* WGBS MM

```{r}
MM_fish_filt$rank <- 1:nrow(MM_fish_filt)
MM_wall_filt$wall_rank <- 1:nrow(MM_wall_filt)
MM_none_filt$rank <- 1:nrow(MM_none_filt)
```
```{r}
MM_fish_filt %>% filter(FDR >= 0.05)
```
```{r}
MM_wall_filt %>% mutate(wall_goseq = GO.wall_seq[match(MM_wall_filt$GOID, GO.wall_seq$category), "wall_rank"],
                   fish = MM_fish_filt[match(MM_wall_filt$GOID, MM_fish_filt$GOID), "rank"]) %>%
  ggplot(aes(y = wall_rank, x = wall_goseq)) +
  geom_point(alpha = 0.3) 
```
```{r}
ggVennDiagram(list(wall_mm = filter(MM_wall_filt, FDR < 0.05)$GOID, wall_goseq = filter(GO.wall_seq, FDR < 0.05)$category))
```


```{r}
MM_wall_filt %>% mutate(none = MM_none_filt[match(MM_wall_filt$GOID, MM_none_filt$GOID), "rank"],
                   fish = MM_fish_filt[match(MM_wall_filt$GOID, MM_fish_filt$GOID), "rank"]) %>%
  ggplot(aes(y = wall_rank, x = none)) +
  geom_point(alpha = 0.3) 
#  geom_vline(xintercept = 453, colour = "red") +
 # geom_hline(yintercept = 1, colour = "blue")

MM_wall_filt %>% mutate(none = MM_none_filt[match(MM_wall_filt$GOID, MM_none_filt$GOID), "rank"],
                   fish = MM_fish_filt[match(MM_wall_filt$GOID, MM_fish_filt$GOID), "rank"]) %>%
  ggplot(aes(y = fish, x = none)) +
  geom_point(alpha = 0.3) 
#  geom_vline(xintercept = 453, colour = "red") +
#  geom_hline(yintercept = 1, colour = "blue")

MM_wall_filt %>% mutate(none = MM_none_filt[match(MM_wall_filt$GOID, MM_none_filt$GOID), "rank"],
                   fish = MM_fish_filt[match(MM_wall_filt$GOID, MM_fish_filt$GOID), "rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish)) +
  geom_point(alpha = 0.3) 
#  geom_vline(xintercept = 453, colour = "red") +
 # geom_hline(yintercept = 1, colour = "blue")
```
```{r}
test_out_filt$de %>% data.frame() %>% filter(`.` == 1)
```
```{r}
mm_array_res <- readRDS("../output/mm_array_res.rds")
```


## change in rank
```{r}
go_entrez2
```
```{r}
MM_wall_filt %>% mutate(none = MM_none_filt[match(MM_wall_filt$GOID, MM_none_filt$GOID), "rank"],
                   fish = MM_fish_filt[match(MM_wall_filt$GOID, MM_fish_filt$GOID), "rank"],
                   mean_cg = go_entrez2[match(MM_wall_filt$GOID, go_entrez2$go), "mean_cg"]) +
  ggplot(aes())
```

```{r}
GO.wall_seq %>% mutate(none = GO.fish_seq[match(GO.wall_seq$category, GO.fish_seq$category), "fish_rank"]) %>%
  ggplot(aes(x = wall_rank, y = fish_seq))
```


```{r}
MM_fish
```

```{r}
plotB
```

```{r}
plotBiasGrouped(biomart_genes, test_anno) + 
```
```{r}
.plotBias(D=test_out_filt$de, bias=as.vector(test_out_filt$freq_genes))
```

```{r}
D <- test_out_filt$de

prior.prob <- bias
  o <- order(bias)
  prior.prob[o] <- limma::tricubeMovingAverage(D[o],span=0.5)
  prior.prob
```

```{r}
mapped_entrez$freq %>% data.frame()
```



```{r}
res <- readRDS("../output/mm_fn_res_num.rds")
res
```

```{r}
compare <- readRDS("../output/go_res_compare.rds")
compare
```
```{r}

```

