---
title: "08_broadGeneSetDatabase"
author: "Caitlin Page"
date: "2025-05-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

```{r}
#renv::install("bioc::KEGGREST")
#renv::install("bioc::msigdb")
```
## MsigDB Analysis

[following vignette](https://bioconductor.org/packages/release/data/experiment/vignettes/msigdb/inst/doc/msigdb.html)


The database is broken down into collections, including GO. Collections are available for human and mouse. It is advised to not run analysis on all gene sets, but on a specific collection.
[Description of collections](https://www.gsea-msigdb.org/gsea/msigdb/human/collections.jsp#C5)
```{r}
library(msigdb)
library(ExperimentHub)
library(GSEABase)
```

```{r}
eh <- ExperimentHub()
query(eh , 'msigdb')
```

```{r}
msigdb.hs <- getMsigdb(org = 'hs', id = 'SYM', version = '7.4')
msigdb.hs
```

```{r}
msigdb.hs
```

data is broken up into chromosome and position
```{r}
example_gs <- msigdb.hs[[1000]]
```

```{r}
geneIds(example_gs)
```
```{r}
geneIdType(example_gs)
```

```{r}
collectionType(example_gs)
```

```{r}
bcCategory(collectionType(example_gs))
```

```{r}
description(example_gs)
```

```{r}
details(example_gs)
```

# summarise info

```{r}
table(sapply(lapply(msigdb.hs, collectionType), bcCategory))
```

```{r}
table(sapply(lapply(msigdb.hs, collectionType), bcSubCategory))
```

```{r}
hist(sapply(lapply(msigdb.hs, geneIds), length),
     main = 'MSigDB signature size distribution',
     xlab = 'Signature size')
```

# subset collection for analysis

```{r}
listCollections(msigdb.hs)
```

```{r}
listSubCollections(msigdb.hs)
```


```{r}
subsetCollection(msigdb.hs, 'h')
```

```{r}
subsetCollection(msigdb.hs, 'c5', 'GO:BP')
```

# prepare for usage
```{r}
library(limma)
```
make expression data
```{r}
allg <- unique(unlist(geneIds(msigdb.hs)))
emat <- matrix(0, nrow = length(allg), ncol = 6)
rownames(emat) <- allg
colnames(emat) <- paste0('sample', 1:6)
head(emat)
```
get collection
```{r}
hallmarks <- subsetCollection(msigdb.hs, 'h')
msigdb_ids <- geneIds(hallmarks)
```
convert gene sets into list of gene indices

```{r}
fry_indices = ids2indices(msigdb_ids, rownames(emat))
fry_indices[1:2]
```

# ok so now what?
how do we do the test?
what did this fry thing do? 
are those gene indices like entrez gene ids?
or is it the position of the gene symbol from the other matrix thing?
```{r}
length(emat)
```

```{r}
fry_indices[[1]][1:5] %>% class()
```
ok this implies to me that it is the index of the gene symbol, not an entrez id (entrez id is character)

but we can get entrez ids with a different function

# implement into function
want something like this
```{r}
mm_out <- readRDS("../output/mm_fn_out.rds")
mm_out$sig.eg[1:3]
mm_out$universe[1:5]
mm_out$freq %>% data.frame() %>% head()
mm_out$equiv %>% data.frame() %>% head()
mm_out$de[1:5]
```
which doesn't even involve gene sets yet
which means that part can stay as it is

all I need is a collection that is a list with the gene sets, and within that is entrez ids

so the important functions to make that happen are...
```{r}
msigdb.hs <- getMsigdb(org = 'hs', id = 'EZID', version = '7.4')
```
basically just this
org can be "hs" or "mm" (human or mouse)
id can be "SYM" or "EZID" (symbol or entrez gene id)
and then version - which is based on version of MSigDb

then there was some weird make expression data stuff which I don't think we need it was weird

want to subset the collection
```{r}
hallmarks <- subsetCollection(gsc = msigdb.hs, collection = 'h')
msigdb_ids <- geneIds(hallmarks)
msigdb_ids$HALLMARK_TNFA_SIGNALING_VIA_NFKB
```
(sub)/collection must be part of list(Sub)Collection or else:
Error in subsetCollection(gsc = msigdb.hs, collection = "a") : 
  all(collection %in% listCollections(gsc)) is not TRUE

all we had to do was add the gene ids to the gene sets
and that's it

change code:
* add in collection parameter for source and subcollection parameter for msigdb
* add appropriate checks
* new small function to make the msigdb object

```{r}
getMsig <- function(species, version = version, collection, subcollection = NULL) {
  species <- species
  msigdb <- msigdb::getMsigdb(org = species, id = 'EZID', version = version)
  hallmarks <- msigdb::subsetCollection(gsc = msigdb, collection = collection, subcollection = subcollection)
  msigdb_ids <- GSEABase::geneIds(hallmarks)
  msigdb_ids
}
getMsig("hs", "7.4", "h")
```
```{r}
listCollections(msigdb.hs)
listSubCollections(msigdb.hs)
```
* incorporate into testing function miss_methyl_b
```{r}
run_gst_seq <- function(dmrs, tested_cpgs,
                        gene_source = c("biomaRt", "ExperimentHub"), gene_feature=c("gene", "promoter"),
                        collection = c("GO", "MsigDb"), subcollection = NULL,
                        method = c("Wallenius", "Fishers", "None"), plot_bias=TRUE, ...) {
  gene_source <- match.arg(tolower(gene_source), c("biomart", "experimenthub"))
  gene_feature <- match.arg(tolower(gene_feature), c("gene", "promoter"))
  genes <- sourceGenes(counts = tested_cpgs, gene_source = gene_source, gene_feature = gene_feature)
  if(nrow(genes) == 0) {
    stop("sourceGenes failed")
  }
  if(nrow(dmrs) > 30000) {
    stop("Please appropriately filter DMRs to most significant") #idk
  }
  anno <- annoGeneDmr(counts=tested_cpgs, dmrs = dmrs, gene_source = gene_source, genes = genes)
  if(nrow(anno) == 0) {
    stop("anno failed")
  }
  out <- run_miss_methyl_a(genes, anno)
  
  collection <- match.arg(tolower(collection), c("go", "msigdb"))
  
  results <- run_miss_methyl_b(out, correct_bias=correct_bias, plot_bias=plot_bias, ...)

  list(res = results, genes = genes, anno = anno, out = out)
}
```

## KEGG
```{r}
?limma::getGeneKEGGLinks()
```

missMethyl does KEGG as well
.getKEGG <- function(){
  GeneID.PathID <- limma::getGeneKEGGLinks(species.KEGG = "hsa", convert = TRUE)
  GeneID.PathID$PathwayID <- gsub("path:", "", GeneID.PathID$PathwayID)
  isna <- rowSums(is.na(GeneID.PathID[, 1:2])) > 0.5
  GeneID.PathID <- GeneID.PathID[!isna, ]
  ID.ID <- paste(GeneID.PathID[, 1], GeneID.PathID[, 2], sep = ".")
  d <- !duplicated(ID.ID)
  GeneID.PathID <- GeneID.PathID[d, ]
  PathID.PathName <- limma::getKEGGPathwayNames(species.KEGG = "hsa", 
                                         remove.qualifier = TRUE)
  #PathID.PathName$PathwayID <- paste0("path:", PathID.PathName$PathwayID)
  GeneID.PathID <- merge(GeneID.PathID, PathID.PathName, by="PathwayID")
  kegg <- tapply(GeneID.PathID$GeneID, GeneID.PathID$PathwayID, list)
  
  list(idList = kegg, idTable = PathID.PathName)
}
