---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Welcome to my research website.

The main analyses conducted uses B cell type and NK cell type samples, 
under the hg19 genome annotation.
* Microarray: EPICv1 platform. [GSE110554](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE110554)
* WGBS [GSE186458](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE186458)

Analysis:

1. Differential methylation methods and comparison

  a) [Microarray](00_differentialMethylationArray.html)
  
  b) Sequencing *tbd 00_differentialMethylationSequencing
  
  c) Compare microarray and sequencing DMRcate results 

2. [Methylation microarray gene set testing analysis](00_microarrayOntologyAnalysis.Rmd)

```{r}
	
#save.image(file='myEnvironment.RData')
```

https://www.techcoil.com/blog/how-to-save-and-load-environment-objects-in-r/#google_vignette
```{r}
#load('myEnvironment.RData')
```





[analysis/00_exploreCpGDistribution]
* plots comparing array and sequencing
** distance between cpgs, annotation to cpg and gene features



